---
title: "Geometry operations"
author: "Xiaochi"
date: "26/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(tmap)
```

# Geometric operations on vector data

## Simplification

```{r}
seine_simp = st_simplify(seine, dTolerance = 2000)  # 2000 m
```

```{r}
object.size(seine)
object.size(seine_simp)
```

```{r}
us_states2163 = st_transform(us_states, 2163)
```

```{r}
us_states_simp1 = st_simplify(us_states2163, dTolerance = 100000)  # 100 km
```

```{r}
# proportion of points to retain (0-1; default 0.05)
us_states2163$AREA = as.numeric(us_states2163$AREA)
us_states_simp2 = rmapshaper::ms_simplify(us_states2163, keep = 0.01,
                                          keep_shapes = TRUE)
```


## Centroids

```{r}
nz_centroid = st_centroid(nz)
seine_centroid = st_centroid(seine)
```

```{r}
nz_pos = st_point_on_surface(nz)
seine_pos = st_point_on_surface(seine)
```


## Buffers

```{r}
seine_buff_5km = st_buffer(seine, dist = 5000)
seine_buff_50km = st_buffer(seine, dist = 50000)
```



## Affine transformations

```{r}
nz_sfc = st_geometry(nz)
```

```{r}
nz_shift = nz_sfc + c(0, 100000)
```

```{r}
nz_centroid_sfc = st_centroid(nz_sfc)
nz_scale = (nz_sfc - nz_centroid_sfc) * 0.5 + nz_centroid_sfc
```

```{r}
rotation = function(a){
  r = a * pi / 180 #degrees to radians
  matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2, ncol = 2)
}

nz_rotate = (nz_sfc - nz_centroid_sfc) * rotation(30) + nz_centroid_sfc
```

```{r}
nz_scale_sf = st_set_geometry(nz, nz_scale)
```


## Clipping

```{r}
b = st_sfc(st_point(c(0, 1)), st_point(c(1, 1))) # create 2 points
b = st_buffer(b, dist = 1) # convert points to circles
plot(b)
text(x = c(-0.5, 1.5), y = 1, labels = c("x", "y")) # add text
```

```{r}
x = b[1]
y = b[2]
x_and_y = st_intersection(x, y)
plot(b)
plot(x_and_y, col = "lightgrey", add = TRUE) # color intersecting area
```

```{r}
bb = st_bbox(st_union(x, y))
box = st_as_sfc(bb)
set.seed(2017)
p = st_sample(x = box, size = 10)
plot(box)
plot(x, add = TRUE)
plot(y, add = TRUE)
plot(p, add = TRUE)
text(x = c(-0.5, 1.5), y = 1, labels = c("x", "y"))
```

```{r}
sel_p_xy = st_intersects(p, x, sparse = FALSE)[, 1] &
  st_intersects(p, y, sparse = FALSE)[, 1]
p_xy1 = p[sel_p_xy]
p_xy2 = p[x_and_y]
identical(p_xy1, p_xy2)
```


## Geometry unions

```{r}
regions = aggregate(x = us_states[, "total_pop_15"], 
                    by = list(us_states$REGION),
                    FUN = sum, 
                    na.rm = TRUE)

regions2 = us_states %>% 
  group_by(REGION) %>%
  summarize(pop = sum(total_pop_15, na.rm = TRUE))
```

```{r}
us_west = us_states[us_states$REGION == "West", ]
us_west_union = st_union(us_west)
```

```{r}
texas = us_states[us_states$NAME == "Texas", ]
texas_union = st_union(us_west_union, texas)
```


## Type transformations

```{r}
multipoint = st_multipoint(matrix(c(1, 3, 5, 1, 3, 1), ncol = 2))
```

```{r}
linestring = st_cast(multipoint, "LINESTRING")
polyg = st_cast(multipoint, "POLYGON")
```

```{r}
multipoint_2 = st_cast(linestring, "MULTIPOINT")
multipoint_3 = st_cast(polyg, "MULTIPOINT")
all.equal(multipoint, multipoint_2, multipoint_3)
```

```{r}
rbind(c(5,2),
      c(1,3),
      c(3,4),
      c(3,2))
```

```{r}
list(
  rbind(c(5,2),
        c(1,3),
        c(3,4),
        c(3,2))
  )
```

```{r}
multilinestring_list = list(matrix(c(1, 4, 5, 3), ncol = 2), 
                            matrix(c(4, 4, 4, 1), ncol = 2),
                            matrix(c(2, 4, 2, 2), ncol = 2))
multilinestring = st_multilinestring((multilinestring_list))
multilinestring_sf = st_sf(geom = st_sfc(multilinestring))
multilinestring_sf
```

```{r}
linestring_sf2 = st_cast(multilinestring_sf, "LINESTRING")
linestring_sf2
```


# Geometric operations on raster data

## Geometric intersections

```{r}
data("elev", package = "spData")
clip = raster(xmn = 0.9, xmx = 1.8, ymn = -0.45, ymx = 0.45,
              res = 0.3, vals = rep(1, 9))
elev[clip, drop = FALSE]
```


## Extent and ortigin

```{r}
data(elev, package = "spData")

elev_2 = extend(elev, c(1, 2), value = 1000)
plot(elev_2)
plot(elev, add = TRUE)
```

```{r}
elev_3 = elev + elev_2
```

```{r}
elev_4 = extend(elev, elev_2)
plot(elev_4)
plot(elev)
```

```{r}
origin(elev_4) # return the raster origin's coordinate
```


```{r}
# change the origin
origin(elev_4) = c(0.25, 0.25)
plot(elev_4)
# returen the coordinate of raster origin
origin(elev_4)
origin(elev)
# and add the original raster
plot(elev, add = TRUE)
```



## Aggregation and disaggregation

```{r}
data("dem", package = "RQGIS")

dem_agg = aggregate(dem, fact = 5, fun = mean)

plot(dem)
plot(dem_agg)
```

```{r}
dem_disagg = disaggregate(dem_agg, fact = 5, method = "")
plot(dem_disagg)
dem_disagg_bilinear = disaggregate(dem_agg, fact = 5, method = "bilinear")
plot(dem_disagg_bilinear)
```


```{r}
identical(dem, dem_disagg)
identical(dem, dem_disagg_bilinear)
```


```{r}
# add 2 rows and columns, i.e. change the extent
dem_agg = extend(dem_agg, 2)
dem_disagg_2 = resample(dem_agg, dem)

```

# Raster-vector interactions

## Raster cropping

```{r}
srtm = raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion = st_read(system.file("vector/zion.gpkg", package = "spDataLarge"))
zion = st_transform(zion, projection(srtm))

plot(srtm)
plot(zion, alpha = 0, add = TRUE)
```

```{r}
srtm_cropped = crop(srtm, zion)
plot(srtm)
plot(srtm_cropped)
```

```{r}
srtm_masked = mask(srtm, zion)
plot(srtm_masked)
```

```{r}
srtm_update0_masked = mask(srtm, zion, updatevalue = 0)
plot(srtm_update0_masked)
```

```{r}
srtm_inv_masked = mask(srtm, zion, inverse = TRUE)
plot(srtm_inv_masked)
```

## Raster extraction

```{r}
data("zion_points", package = "spDataLarge")
zion_points$elevation = raster::extract(srtm, zion_points)
```

```{r}
raster::extract(srtm, zion_points, buffer = 100)
```


```{r}
zion_transect = cbind(c(-113.2, -112.9), c(37.45, 37.2)) %>%
  st_linestring() %>% 
  st_sfc(crs = projection(srtm)) %>% 
  st_sf()
```

```{r}
transect = raster::extract(srtm, zion_transect, 
                           along = TRUE, cellnumbers = TRUE)
```

```{r}
transect_df = purrr::map_dfr(transect, as_data_frame, .id = "ID")
transect_coords = xyFromCell(srtm, transect_df$cell)
pair_dist = geosphere::distGeo(transect_coords)[-nrow(transect_coords)]
transect_df$dist = c(0, cumsum(pair_dist)) 
```

```{r}
zion_srtm_values = raster::extract(x = srtm, y = zion, df = TRUE) 
```

```{r}
group_by(zion_srtm_values, ID) %>% 
  summarize_at(vars(srtm), list(~min(.), ~mean(.), ~max(.)))
```

```{r}
zion_nlcd = raster::extract(nlcd, zion, df = TRUE, factors = TRUE) 
dplyr::select(zion_nlcd, ID, levels) %>% 
  tidyr::gather(key, value, -ID) %>%
  group_by(ID, key, value) %>%
  tally() %>% 
  tidyr::spread(value, n, fill = 0)
```


## Rasterization

```{r}
cycle_hire_osm_projected = st_transform(cycle_hire_osm, 27700)

raster_template = raster(
  extent(cycle_hire_osm_projected), 
  resolution = 1000,
  crs = st_crs(cycle_hire_osm_projected)$proj4string
  )

plot(cycle_hire_osm_projected)
```

```{r}
ch_raster1 = rasterize(cycle_hire_osm_projected, raster_template, field = 1)
ch_raster1[,]
plot(ch_raster1)
```

```{r}
ch_raster2 = rasterize(cycle_hire_osm_projected, raster_template, 
                       field = 1, fun = "count")
ch_raster2[,]
plot(ch_raster2)
```

```{r}
ch_raster3 = rasterize(cycle_hire_osm_projected, raster_template, 
                       field = "capacity", fun = sum)
ch_raster3[,]
values(ch_raster3)
getValues(ch_raster3)

stack(ch_raster1, ch_raster2, ch_raster3)[1]

r_stack = stack(ch_raster1, ch_raster2, ch_raster3)
names(r_stack) = c("ch_raster1", "ch_raster2", "ch_raster3")
raster::subset(r_stack, "ch_raster1")
r_stack[["ch_raster1"]]
r_stack$ch_raster1

ch_raster3[1] = NA


plot(ch_raster3)
```


```{r}
california = dplyr::filter(us_states, NAME == "California")
plot(california)

california_borders = st_cast(california, "MULTILINESTRING")
plot(california_borders)

raster_template2 = raster(
  extent(california), 
  resolution = 0.5,
  crs = st_crs(california)$proj4string
  )
```

```{r}
california_raster1 = rasterize(california_borders, raster_template2)
california_raster1[,]
plot(california_raster1)
```


## Spatial vectorization

```{r}
plot(elev)
elev_point = rasterToPoints(elev, spatial = TRUE) %>% 
  st_as_sf()
plot(elev_point)
```

```{r}
data(dem, package = "RQGIS")
cl = rasterToContour(dem)
plot(dem, axes = FALSE)
plot(cl, add = TRUE)
```

```{r}
# create hillshade
hs = hillShade(slope = terrain(dem, "slope"), aspect = terrain(dem, "aspect"))
plot(hs, col = gray(0:100 / 100), legend = FALSE)
# overlay with DEM
plot(dem, 
     col = terrain.colors(25), 
     alpha = 0.5, 
     legend = FALSE, 
     add = TRUE)
# add contour lines
contour(dem, col = "white", add = TRUE)
```

```{r}
grain_poly = rasterToPolygons(grain) %>% 
  st_as_sf()

plot(grain)
plot(grain_poly)

grain_poly2 = grain_poly %>% 
  group_by(layer) %>%
  summarize()

plot(grain_poly2)
```

