---
title: "Scripts, algorithms and functions"
author: "Xiaochi"
date: "29/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scripts

```{r}
source("Hello Geocomputaion with R.R")
```

```{r}
poly_mat = cbind(
  x = c(0, 0, 9, 9, 0),
  y = c(0, 9, 9, 0, 0)
)

source("https://git.io/10-centroid-alg.R") # short url
```


# Geometric algorithms

```{r}
# generate a simple matrix representation of a polygon:
x_coords = c(10, 0, 0, 12, 20, 10)
y_coords = c(0, 0, 10, 20, 15, 0)
poly_mat = cbind(x_coords, y_coords)
plot(poly_mat)
```

```{r}
# create a point representing the origin:
Origin = poly_mat[1, ]
# create 'triangle matrix':
T1 = rbind(Origin, poly_mat[2:3, ], Origin)
plot(T1)
# find centroid (drop = FALSE preserves classes, resulting in a matrix):
C1 = (T1[1, , drop = FALSE] + T1[2, , drop = FALSE] + T1[3, , drop = FALSE]) / 3
```

```{r}
# calculate the area of the triangle represented by matrix T1:
abs(
  T1[1, 1] * (T1[2, 2] - T1[3, 2]) + 
  T1[2, 1] * (T1[3, 2] - T1[1, 2]) + 
  T1[3, 1] * (T1[1, 2] - T1[2, 2]) ) / 2
```

```{r}
i = 2:(nrow(poly_mat) - 2)
T_all = lapply(i, function(x) {
  rbind(Origin, poly_mat[x:(x + 1), ], Origin)
})

C_list = lapply(T_all,  function(x) (x[1, ] + x[2, ] + x[3, ]) / 3)
C = do.call(rbind, C_list)

A = vapply(T_all, function(x) {
  abs(x[1, 1] * (x[2, 2] - x[3, 2]) +
        x[2, 1] * (x[3, 2] - x[1, 2]) +
        x[3, 1] * (x[1, 2] - x[2, 2]) ) / 2
  }, FUN.VALUE = double(1))
```

# Functions

```{r}
t_centroid = function(x) {
  (x[1, ] + x[2, ] + x[3, ]) / 3
}
```

```{r}
t_centroid(T1)
```

```{r}
t_area = function(x) {
  abs(
    x[1, 1] * (x[2, 2] - x[3, 2]) +
    x[2, 1] * (x[3, 2] - x[1, 2]) +
    x[3, 1] * (x[1, 2] - x[2, 2])
  ) / 2
}
```

```{r}
t_area(T1)
```

```{r}
t_new = cbind(x = c(0, 3, 3, 0),
              y = c(0, 0, 1, 0))
t_area(t_new)
```

```{r}
poly_centroid = function(x) {
  i = 2:(nrow(x) - 2)
  T_all = lapply(i, function(x) {
    rbind(Origin, poly_mat[x:(x + 1), ], Origin)
  })
  C_list = lapply(T_all, t_centroid)
  C = do.call(rbind, C_list)
  A = vapply(T_all, t_area, FUN.VALUE = double(1))
  c(weighted.mean(C[, 1], A), weighted.mean(C[, 2], A))
}
```

```{r}
poly_centroid(poly_mat)
```

```{r}
poly_centroid_sfg = function(x) {
  centroid_coords = poly_centroid(x)
  sf::st_point(centroid_coords)
}
```

```{r}
poly_sfc = sf::st_polygon(list(poly_mat))
identical(poly_centroid_sfg(poly_mat), sf::st_centroid(poly_sfc))
```

# Programming

Aim: take a matrix representing a convex polygon, return its centroid, demonstrate how algorithms work

Pre-requisite: an input object named poly_mat with 2 columns representing vertices of a polygon, with 1st and last rows identical:

```{r}
if(!exists("poly_mat")) {
  message("No poly_mat object provided, creating object representing a 9 by 9 square")
  poly_mat = cbind(
    x = c(0, 0, 9, 9, 0),
    y = c(0, 9, 9, 0, 0)
  )
}
```

Step 1: create sub-triangles, set-up

```{r}
Origin = poly_mat[1, ] # create a point representing the origin
i = 2:(nrow(poly_mat) - 2)
T_all = lapply(i, function(x) {
  rbind(Origin, poly_mat[x:(x + 1), ], Origin)
})
```

Step 2: calculate triangle centroids

```{r}
C_list = lapply(T_all,  function(x) (x[1, ] + x[2, ] + x[3, ]) / 3)
C = do.call(rbind, C_list)
```

Step 3: calculate triangle areas

```{r}
A = vapply(T_all, function(x) {
  abs(
    x[1, 1] * (x[2, 2] - x[3, 2]) +
    x[2, 1] * (x[3, 2] - x[1, 2]) +
    x[3, 1] * (x[1, 2] - x[2, 2]) ) / 
    2}, 
  FUN.VALUE = double(1))
```

Step 4: calculate area-weighted centroid average

```{r}
poly_area = sum(A)
print(paste0("The area is: ", poly_area))
poly_centroid = c(weighted.mean(C[, 1], A), weighted.mean(C[, 2], A))
```

Step 5: output results

```{r}
print(paste0(
  "The coordinates of the centroid are: ",
  round(poly_centroid[1], 2),
  ", ",
  round(poly_centroid[2], 2)
))
```

