---
title: "Graphics for Communication"
author: "Xiaochi"
date: "06/09/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(tidyverse)
```

# Label

labs( ): add labels

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov"
  )
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```


```{r}
df <- tibble(
  x = runif(10),
  y = runif(10)
)
ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + beta + frac(delta, theta))
  )
```



* title: the purpose of a plot title is to summarise the main finding. Avoid titles that just describe what the plot is, e.g. “A scatterplot of engine displacement vs. fuel economy”.

* subtitle: adds additional detail in a smaller font beneath the title.

* caption: adds text at the bottom right of the plot, often used to describe the source of the data.

* x

* y

* colour


```{r}
df <- tibble(
  x = runif(10),
  y = runif (10)
)

ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + bete + frac(delta, theta))
  )
```

# Annotations

* geom_text( )
    * label
    * data: have a tibble that provides labels

```{r}
best_in_class <- mpg %>% 
  group_by(class) %>% 
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_text(aes(label = model), data = best_in_class)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
```

```{r}
class_avg <- mpg %>% 
  group_by(class) %>% 
  summarise(
    displ = median(displ),
    hwy = median(hwy)
  )

ggplot(mpg, aes(displ, hwy, colour = class)) +
  ggrepel::geom_label_repel(aes(label = class), data = class_avg,
                            size = 6, label.size = 0, segment.color = NA) +
  geom_point() +
  theme(legend.position = "none")
```

```{r}
label <- mpg %>% 
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \n related to decreasing fuel economy."
  )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```

```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```

```{r}
"Increasing engine size is related to decreasing fuel economy." %>% 
  stringr::str_wrap(width = 40) %>% 
  writeLines()
```

* Use geom_hline() and geom_vline() to add reference lines. I often make them thick (size = 2) and white (colour = white), and draw them underneath the primary data layer. That makes them easy to see, without drawing attention away from the data.

* Use geom_rect() to draw a rectangle around points of interest. The boundaries of the rectangle are defined by aesthetics xmin, xmax, ymin, ymax.

* Use geom_segment() with the arrow argument to draw attention to a point with an arrow. Use aesthetics x and y to define the starting location, and xend and yend to define the end location.

# Scales

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete()
```

* the naming scheme for scales
    * scale
    * _
    * the name of the aesthetic
    * _
    * the name of the scale

* the default scale
    * continuous
    * discrete
    * datetime
    * date

## Axis ticks and legend keys

the ticks on the axes.

the keys on the legend.

Breaks controls the position of the ticks, or the values associated with the keys. 

Labels controls the text label associated with each tick/key.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

Collectively axes and legends are called guides.

*  Axes are used for x and y aesthetics
*  Legends are used for everything else.

```{r}
presidential %>% 
  mutate(id = 33 + row_number()) %>% 
  ggplot(aes(start, id)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y'")
```

* date_labels takes a format specification, in the same form as parse_datetime().

* date_breaks (not shown here), takes a string like “2 days” or “1 month”.

## Legend layout

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
base
```

```{r}
base + theme(legend.position = "left")
```

```{r}
base + theme(legend.position = "top")
```


```{r}
base + theme(legend.position = "bottom")
```
```{r}
base + theme(legend.position = "right") # the default
```

```{r}
base + theme(legend.position = "none")
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))
```

## Replacing a scale

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_color_brewer(palette = "Set1")
  
```


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_color_brewer(palette = "Set1")
```

```{r}
presidential %>% 
  mutate(id = 33 + row_number()) %>% 
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```

Note that all colour scales come in two variety: scale_colour_x( ) and scale_fill_x( ) for the colour and fill aesthetics respectively (the colour scales are available in both UK and US spellings).

# Zooming

There are three ways to control the plot limits: 

1. Adjusting what data are plotted
2. Setting the limits in each scale
3. Setting xlim and ylim in coord_cartesian( )

```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), 
                  ylim = c(10, 30))
```

```{r}
mpg %>% 
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```

```{r}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point()

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point()
```

```{r}
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_color_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
```

# Themes

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE)
```

# Saving your plots

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()
```

```{r}
ggsave("../fig_output/my_plot.pdf")
```


# Color

```{r}
ToothGrowth
mtcars
```

```{r}
# Convert dose and cyl columns from numeric to factor variables
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
mtcars$cyl <- as.factor(mtcars$cyl)
```

```{r}
# Box plot
ggplot(ToothGrowth, aes(x=dose, y=len)) +geom_boxplot()
# scatter plot
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()
```

```{r}
# box plot
ggplot(ToothGrowth, aes(x=dose, y=len)) +
  geom_boxplot(fill='#A4A4A4', color="darkred")
# scatter plot
ggplot(mtcars, aes(x=wt, y=mpg)) + 
  geom_point(color='darkblue')
```

```{r}
# Box plot
bp<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) +
  geom_boxplot()
bp
# Scatter plot
sp<-ggplot(mtcars, aes(x=wt, y=mpg, color=cyl)) + geom_point()
sp
```

```{r}
# Box plot
bp + scale_fill_hue(l=40, c=35)
# Scatter plot
sp + scale_color_hue(l=40, c=35)
```


```{r}
# Box plot
bp + scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
# Scatter plot
sp + scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

```{r}
# Box plot
bp + scale_fill_manual(breaks = c("2", "1", "0.5"), 
                       values=c("red", "blue", "green"))
# Scatter plot
sp + scale_color_manual(breaks = c("8", "6", "4"),
                        values=c("red", "blue", "green"))
```

```{r}
# Box plot
bp + scale_fill_brewer(palette="Dark2")
# Scatter plot
sp + scale_color_brewer(palette="Dark2")
```

```{r}
# Box plot
bp + scale_fill_grey() + theme_classic()
# Scatter plot
sp + scale_color_grey() + theme_classic()
```

```{r}
# Box plot
bp + scale_fill_grey(start=0.8, end=0.2) + theme_classic()
# Scatter plot
sp + scale_color_grey(start=0.8, end=0.2) + theme_classic()
```

```{r}
# Color by qsec values
sp2<-ggplot(mtcars, aes(x=wt, y=mpg, color=qsec)) + geom_point()
sp2
```

```{r}
# Change the low and high colors
# Sequential color scheme
sp2+scale_color_gradient(low="blue", high="red")
```

```{r}
# Diverging color scheme
mid<-mean(mtcars$qsec)
sp2+scale_color_gradient2(midpoint=mid, low="blue", mid="white",
                     high="red", space ="Lab" )
```

```{r}
set.seed(1234)
x <- rnorm(200)
# Histogram
hp<-qplot(x =x, fill=..count.., geom="histogram") 
hp
# Sequential color scheme
hp+scale_fill_gradient(low="blue", high="red")
```


```{r}
# Scatter plot
# Color points by the mpg variable
sp3<-ggplot(mtcars, aes(x=wt, y=mpg, color=mpg)) + geom_point()
sp3
```

```{r}
# Gradient between n colors
sp3+scale_color_gradientn(colours = rainbow(5))
```

# SF

```{r}
library(tidyverse)
library(sf)
system.file("gpkg/nc.gpkg", package="sf") %>% read_sf() %>% 
    st_transform(32119) -> nc.32119
```

```{r}
ggplot() + 
  geom_sf(data = nc.32119) 
```

```{r}
ggplot(nc.32119) + geom_sf()
```

```{r}
ggplot() + 
  geom_sf(data = nc.32119) + 
  theme_void() +
  theme(panel.grid.major = element_line(color = "white"))
```

```{r}
ggplot() + 
  geom_sf(data = nc.32119) + 
  aes(fill = BIR74) +
  scale_fill_gradientn(colors = viridis::viridis(20))
```

```{r}
ggplot() + 
  geom_sf(data = nc.32119) + 
  geom_sf(data = st_centroid(nc.32119))
```

```{r}
nc <- sf::st_read(system.file("gpkg/nc.gpkg", package = "sf"), quiet = TRUE)
nc_3857 <- sf::st_transform(nc, "+init=epsg:3857")
ggplot() + 
    geom_sf(data = nc_3857[1:3, ], aes(fill = AREA)) + 
    geom_sf_label(data = nc_3857[1:3, ], aes(label = NAME))
```

```{r}
library(stars)
library(ggplot2)
library(viridis)
#> Loading required package: viridisLite
system.file("tif/L7_ETMs.tif", package = "stars") %>% read_stars() -> x

g = ggplot() + 
    coord_equal() + 
    scale_fill_viridis() + 
    theme_void() +
    scale_x_discrete(expand=c(0,0)) +
    scale_y_discrete(expand=c(0,0))

g + geom_stars(data = x) + 
    facet_wrap(~band)
```

```{r}
g + geom_stars(data = x, downsample = c(10,10,1)) + 
    facet_wrap(~band)
```

```{r}
library(spacetime)
data(air) # this loads several datasets in .GlobalEnv
d = st_dimensions(station = st_as_sfc(stations), time = dates)
aq = st_as_stars(list(PM10 = air), dimensions = d)
# ggplot() + geom_stars(data = aq[,,3000])
aq.sf = st_as_sf(aq[,,3000], long=TRUE)
ggplot() + 
   geom_sf(data = st_as_sf(DE_NUTS1)) + 
   geom_sf(data = aq.sf, mapping = aes(col = PM10)) + 
   ggtitle(aq.sf$time[1])
```

# Mix Multiple Graphs

```{r}
library(ggpubr)

# ToothGrowth
data("ToothGrowth")
ToothGrowth

# mtcars 
data("mtcars")
mtcars$name <- rownames(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
```

```{r}
# Box plot (bp)
bxp <- ggboxplot(ToothGrowth, x = "dose", y = "len",
                 color = "dose", palette = "jco")
bxp
```

```{r}
# Dot plot (dp)
dp <- ggdotplot(ToothGrowth, x = "dose", y = "len",
                 color = "dose", palette = "jco", binwidth = 1)
dp
```

```{r}
# Bar plot (bp)
bp <- ggbarplot(mtcars, x = "name", y = "mpg",
          fill = "cyl",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",           # Sort the value in ascending order
          sort.by.groups = TRUE,      # Sort inside each group
          x.text.angle = 90           # Rotate vertically x axis texts
          )
bp
```

```{r}
# Scatter plots (sp)
sp <- ggscatter(mtcars, x = "wt", y = "mpg",
                add = "reg.line",               # Add regression line
                conf.int = TRUE,                # Add confidence interval
                color = "cyl", palette = "jco", # Color by groups "cyl"
                shape = "cyl"                   # Change point shape by groups "cyl"
                )+
  stat_cor(aes(color = cyl), label.x = 3)       # Add correlation coefficient
sp
```

## arrange

```{r}
ggarrange(bxp, dp, bp + rremove("x.text"), 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)
```

```{r}
library("cowplot")
plot_grid(bxp, dp, bp + rremove("x.text"), 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)
```

```{r}
library("gridExtra")
grid.arrange(bxp, dp, bp + rremove("x.text"), 
             ncol = 2, nrow = 2)
```

## annotate


```{r fig.width=4}
figure <- ggarrange(sp, bp + font("x.text", size = 10),
                    ncol = 1, nrow = 2)
annotate_figure(figure,
                top = text_grob("Visualizing mpg", 
                                color = "red", face = "bold", size = 14),
                bottom = text_grob("Data source: \n mtcars data set", 
                                   color = "blue", hjust = 1, x = 1, 
                                   face = "italic", size = 10),
                left = text_grob("Figure arranged using ggpubr", 
                                 color = "green", rot = 90),
                right = "I'm done, thanks :-)!",
                fig.lab = "Figure 1", fig.lab.face = "bold"
                )
```

## align plot panels

```{r}
# Fit survival curves
library(survival)
fit <- survfit( Surv(time, status) ~ adhere, data = colon )
# Plot survival curves
library(survminer)
ggsurv <- ggsurvplot(fit, data = colon, 
                     palette = "jco",                              # jco palette
                     pval = TRUE, pval.coord = c(500, 0.4),        # Add p-value
                     risk.table = TRUE                            # Add risk table
                     )
names(ggsurv)
```

```{r fig.width=3}
ggarrange(ggsurv$plot, ggsurv$table, heights = c(2, 0.7),
          ncol = 1, nrow = 2)
```

```{r fig.width=3}
ggarrange(ggsurv$plot, ggsurv$table, heights = c(2, 0.7),
          ncol = 1, nrow = 2, align = "v")
```

## column row span

```{r fig.width=3}
ggarrange(
  # First row with scatter plot 
  sp,
  # Second row with box and dot plots
  ggarrange(bxp, dp, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  # Labels of the scatter plot
  labels = "A") 
```

```{r fig.width=4}
library("cowplot")

ggdraw() +
  draw_plot(bxp, x = 0, y = .5, width = .5, height = .5) +
  draw_plot(dp, x = .5, y = .5, width = .5, height = .5) +
  draw_plot(bp, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot_label(label = c("A", "B", "C"), size = 16,
                  x = c(0, 0.5, 0), y = c(1, 1, 0.5))
```

```{r fig.width=4}
library("gridExtra")

grid.arrange(
  # First row with one plot spaning over 2 columns
  sp,
  # Second row with 2 plots in 2 different columns
  arrangeGrob(bxp, dp, ncol = 2),
  # Number of rows
  nrow = 2) 
```

```{r fig.width=4}
grid.arrange(bp,                                    # bar plot spaning two columns
             bxp, sp,                               # box plot and scatter plot
             ncol = 2, nrow = 2, 
             layout_matrix = rbind(c(1,1), c(2,3)))
```

```{r fig.width=4}
library("gridExtra")
library("cowplot")
# Arrange plots using arrangeGrob
# returns a gtable (gt)
gt <- arrangeGrob(
  # bar plot spaning two columns
  bp, 
  # box plot and scatter plot
  bxp, sp,                               
  ncol = 2, nrow = 2, 
  layout_matrix = rbind(c(1,1), c(2,3)))

# Add labels to the arranged plots
p <- as_ggplot(gt) + # transform to a ggplot
  draw_plot_label(label = c("A", "B", "C"), size = 15,
                  x = c(0, 0, 0.5), y = c(1, 0.5, 0.5)) # Add labels

p
```

```{r fig.width=4}
library(grid)
# Move to a new page
grid.newpage()
# Create layout : nrow = 3, ncol = 2
pushViewport(viewport(layout = grid.layout(nrow = 3, ncol = 2)))
# A helper function to define a region on the layout
define_region <- function(row, col){
  viewport(layout.pos.row = row, layout.pos.col = col)
} 
# Arrange the plots
print(sp, vp = define_region(row = 1, col = 1:2))   # Span over two columns
print(bxp, vp = define_region(row = 2, col = 1))
print(dp, vp = define_region(row = 2, col = 2))
print(bp + rremove("x.text"), vp = define_region(row = 3, col = 1:2))
```

## common legend

```{r}
ggarrange(bxp, dp, labels = c("A", "B"),
          common.legend = TRUE, legend = "bottom")
```

## scater + density

```{r}
# Scatter plot colored by groups ("Species")
sp <- ggscatter(iris, x = "Sepal.Length", y = "Sepal.Width",
                color = "Species", palette = "jco",
                size = 3, alpha = 0.6) +
  border()

# Marginal density plot of x (top panel) and y (right panel)
xplot <- ggdensity(iris, "Sepal.Length", fill = "Species",
                   palette = "jco")
yplot <- ggdensity(iris, "Sepal.Width", fill = "Species", 
                   palette = "jco") +
  rotate()

# Cleaning the plots
yplot <- yplot + clean_theme() 
xplot <- xplot + clean_theme()

# Arranging the plot
ggarrange(xplot, NULL, sp, yplot, 
          ncol = 2, nrow = 2,  align = "hv", 
          widths = c(2, 1), heights = c(1, 2),
          common.legend = TRUE)
```

