---
title: "Iteration"
author: "Xiaochi"
date: "20/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(tidyverse)
```

reducing duplication benefits:

1. It's easier to see the intent of your code.

2. It's easier to respond to changes in requirements.

3. You're likely have fewer bugs.

reducing duplication tools:

* function
  + identifying repeated patterns of cod
  + extract them out into independent pieces

* iteration
  + imperative programming: for loops, while loops
  + functional programming: each common for loop pattern gets its own function.

# For loops

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df
```


```{r}
median(df$a)
median(df$b)
median(df$c)
median(df$d)
```

```{r}
output <- vector("double",ncol(df)) # 1. output
for (i in seq_along(df)) {          # 2. sequence
  output[[i]] <- median(df[[i]])    # 3. body
}
output
```

loops:

1. output
  + to be efficient
  + vector()
      + the type of the vector
      + the length of the vector

2. sequence
  + i as a pronoun, like "it"
  + seq_along(), a safe version of 1:length()
  
```{r}
y <- vector("double", 0)
length(y)
1:length(y)
seq_along(y)
```


3. The body

# For loop variation

There are four variations on the basic theme of the for loop:

1. Modifying an existing object, instead of creating a new object.
2. Looping over names or values, instead of indices.
3. Handling outputs of unknown length.
4. Handling sequences of unknown length.

## Modifying an existing object

to rescale every column in a data frame:

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df
```

```{r}
rescale01 <- function(x) {
  # range returns a vector containing the minimum and maximum of all the given arguments.
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

rescale01(df$a)
rescale01(df$b)
rescale01(df$c)
rescale01(df$d)
```

To solve this with a for loop we again think about the three components:

1. Output: the same as the input!

2. Sequence: we can think about a data frame as a list of columns, so we can iterate over each column with seq_along(df).

3. Body: apply rescale01().

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df
```

```{r}
for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}
df
```

Typically you’ll be modifying a list or data frame with this sort of loop, so remember to use [[, not [. You might have spotted that I used [[ in all my for loops: I think it’s better to use [[ even for atomic vectors because it makes it clear that I want to work with a single element.

## Looping patterns

1. loop over the numeric indices
  + for (i in seq_along(xs))
  + x[[i]]

2. loop over the elements
  + for (x in xs)

3. loop over the names
  + for (nm in names(xs))
  + x[[nm]]

Iteration over the numeric indices is the most general form, because given the position you can extract both the name and the value:

## Unknown output length

simulate some random vectors of random lengths

```{r}
# different means
means <- c(0, 1, 2)

output <- double()

for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)
```

But this is not very efficient because in each iteration, R has to copy all the data from the previous iterations.

A better solution to save the results in a list, and then combine into a single vector after the loop is done:

```{r}
out <- vector("list",length(means))
for(i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}
str(out)
str(unlist(out))
```

This pattern occurs in other places too:

* generating a long string:
  + save the output in a character vector
  + combine that vector into a single string
    + paste(output, collapse = "")

* generatign a big data frame
  + save the output in a list
  + combine the output into a single data frame
    + dplyr::bind_rows(output)

## Unknown sequence length

A while loop

* condition
* body

Here’s how we could use a while loop to find how many tries it takes to get three heads in a row:

```{r}
flip <- function() sample(c("T", "H"), 1)

flips <- 0
nheads <- 0

while (nheads < 3) {
  if (flip() == "H") {
    nheads <- nheads + 1
  } else {
    nheads <- 0
  }
  flips <- flips + 1
}

flips
```

# For loops vs. functions

For loops are not as important in R as they are in other languages because R is a functional programming language. This means that it’s possible to wrap up for loops in a function, and call that function instead of using the for loop directly.

To see why this is important, consider (again) this simple data frame:

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df
```

Imagine you want to compute the mean of every column. You could do that with a for loop:

```{r}
output <- vector("double",length(df))
for(i in seq_along(df)) {
  output[[i]] <- mean(df[[i]])
}
output
```

You realise that you’re going to want to compute the means of every column pretty frequently, so you extract it out into a function:

```{r}
col_mean <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- mean(df[[i]])
  }
  output
}
```

But then you think it’d also be helpful to be able to compute the median, and the standard deviation, so you copy and paste your col_mean() function and replace the mean() with median() and sd():

```{r}
col_median <- function(df) {
  output <- vector("double", length(df))
  for(i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}

col_sd <- function(df) {
  output <- vector("double", length(df))
  for(i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}
```

Uh oh! You’ve copied-and-pasted this code twice, so it’s time to think about how to generalise it. Notice that most of this code is for-loop boilerplate and it’s hard to see the one thing (mean(), median(), sd()) that is different between the functions.

What would you do if you saw a set of functions like this:

```{r}
f1 <- function(x) abs(x - mean(x)) ^ 1
f2 <- function(x) abs(x - mean(x)) ^ 2
f3 <- function(x) abs(x - mean(x)) ^ 3
```

Hopefully, you’d notice that there’s a lot of duplication, and extract it out into an additional argument:

```{r}
f <- function(x, i) abs(x - mean(x)) ^ i
```

You’ve reduced the chance of bugs (because you now have 1/3 of the original code), and made it easy to generalise to new situations.

We can do exactly the same thing with col_mean(), col_median() and col_sd() by adding an argument that supplies the function to apply to each column:

```{r}
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}

col_summary(df, median)

col_summary(df, mean)
```

The idea of passing a function to another function is extremely powerful idea, and it’s one of the behaviours that makes R a functional programming language. 

The goal of using purrr functions instead of for loops is to allow you break common list manipulation challenges into independent pieces:

1. How can you solve the problem for a single element of the list? Once you’ve solved that problem, purrr takes care of generalising your solution to every element in the list.

2. If you’re solving a complex problem, how can you break it down into bite-sized pieces that allow you to advance one small step towards a solution? With purrr, you get lots of small pieces that you can compose together with the pipe.

This structure makes it easier to solve new problems. It also makes it easier to understand your solutions to old problems when you re-read your old code.

# The map functions

The pattern of 

1. looping over a vector, 
2. doing something to each element
3. saving the results 

is so common that the purrr package provides a family of functions to do it for you. There is one function for each type of output:

* map() makes a list.
* map_lgl() makes a logical vector.
* map_int() makes an integer vector.
* map_dbl() makes a double vector.
* map_chr() makes a character vector.

Each function 

* takes a vector as input, 
* applies a function to each piece, 
* returns a new vector that’s the same length (and has the same names) as the input. The type of the vector is determined by the suffix to the map function.


summary functions

```{r}
df
```

```{r}
map_dbl(df, mean)
```

```{r}
map_dbl(df, median)
```

```{r}
map_dbl(df, sd)
```

focusing on operation.

```{r}
df %>% map_dbl(mean)
```

```{r}
df %>% map_dbl(median)
```

```{r}
df %>% map_dbl(sd)
```

map_*()

* The second argumnet, .f, the function to apply
  + a furmula
  + a caracter vector
  + an integer vector.

* map_*() uses ... to pass along additional arguments to .f each time it's called.

```{r}
map_dbl(df, mean, trim = 0.5)
```

* The map functions also preserve names:

```{r}
z <- list(x = 1:3, y = 4:5)
str(z)
map_int(z, length)
```

## Shortcuts

fit a linear model to each group in a dataset.

1. splits up the mtcars dataset into three pieces (one for each value of cylinder)

```{r}
mtcars %>% 
  split(.$cyl)
```

2. fits the same linear model ot each piece:

```{r}
mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))
```



* There is an anonymous function. purrr provides a convenient shortcut: a one-sided formula:

* . as a pronoun: it refers to the current list element (in the same way that i refereed to the current index in the for loop).

```{r}
mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))
```

many models: extract a summary statistic liek the R2.

```{r}
many_models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))

many_models %>% 
  map(summary)
```

```{r}
many_models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)
```

But extracting named components is a common operation, so purrr provides an even shorter shortcut: you can use a string.

```{r}
many_models %>% 
  map(summary) %>% 
  map_dbl("r.squared")
```

use an integer to select elements by position:

```{r}
x <- list(list(1, 2, 3), list(4, 5, 6), list(7, 8, 9))
str(x)
```

```{r}
x %>% map_dbl(2)
```

## Base R

If you’re familiar with the apply family of functions in base R, you might have noticed some similarities with the purrr functions:

* lapply() is basically identical to map(), except that map() is consistent with all the other functions in purrr, and you can use the shortcuts for .f.

* Base sapply() is a wrapper around lapply() that automatically simplifies the output. This is useful for interactive work but is problematic in a function because you never know what sort of output you’ll get:

```{r}
threshold <- function(x, cutoff = 0.8) x[x > cutoff]
```


```{r}
x1 <- list(
  c(0.27, 0.37, 0.57, 0.91, 0.20),
  c(0.90, 0.94, 0.66, 0.63, 0.06), 
  c(0.21, 0.18, 0.69, 0.38, 0.77)
)
str(x1)
```

```{r}
x1 %>% lapply(threshold) %>% str()
```

```{r}
x1 %>% sapply(threshold) %>% str()
```

```{r}
x2 <- list(
  c(0.50, 0.72, 0.99, 0.38, 0.78), 
  c(0.93, 0.21, 0.65, 0.13, 0.27), 
  c(0.39, 0.01, 0.38, 0.87, 0.34)
)
str(x2)
```

```{r}
x2 %>% lapply(threshold) %>% str()
```

```{r}
x2 %>% sapply(threshold) %>% str()
```

* vapply() is a safe alternative to sapply() because you supply an additional argument that defines the type. The only problem with vapply() is that it’s a lot of typing: vapply(df, is.numeric, logical(1)) is equivalent to map_lgl(df, is.numeric). One advantage of vapply() over purrr’s map functions is that it can also produce matrices — the map functions only ever produce vectors.


# Dealing with failure

safely( ). safely( ) is an adverb: it takes a function (a verb) and returns a modified version. In this case, the modified function will never throw an error. Instead, it always returns a list with two elements:

1. result is the original result. If there was an error, this will be NULL.

2. error is an error object. If the operation was successful, this will be NULL.

Let’s illustrate this with a simple example: log():

```{r}
safe_log <- safely(log)
str(safe_log(10))
```

```{r}
safe_log("a")
str(safe_log("a"))
```

* When the function succeeds, the result element contains the result and the error element is NULL. 
* When the function fails, the result element is NULL and the error element contains an error object.

safely() is designed to work with map:

```{r}
x <- list(1, 10, "a")
str(x)
```

```{r}
y <- x %>% map(safely(log))
str(y)
```

This would be easier to work with if we had two lists: one of all the errors and one of all the output. That’s easy to get with purrr::transpose():

```{r}
y <- y %>% transpose()
str(y)
```

It’s up to you how to deal with the errors,

* look at the values of x where y is an error
* or work with the values of y that are ok

```{r}
y$error
is_ok <- y$error %>% map_lgl(is_null)
is_ok
x[!is_ok]
y$result[is_ok] %>% flatten_dbl()
```

Purrr provides two other useful adverbs:

* Like safely( ), possibly( ) always succeeds. It’s simpler than safely( ), because you give it a default value to return when there is an error.

```{r}
x <- list(1, 10, "a")
str(x)
x %>% map_dbl(possibly(log,NA_real_))
```

* quietly( ) performs a similar role to safely( ), but instead of capturing errors, it captures printed output, messages, and warnings:

```{r}
x <- list(1, -1)
str(x)
x %>% map(quietly(log)) %>% str()
```

# Mapping over multiple arguments

So far we’ve mapped along a single input. But often you have multiple related inputs that you need iterate along in parallel. That’s the job of the map2( ) and pmap( ) functions. 

For example, imagine you want to simulate some random normals with different means. You know how to do that with map():

```{r}
mu <- list(5, 10, -3)
str(mu)
```

```{r}
mu %>% 
  map(rnorm, n = 5) %>% 
  str()
```

What if you also want to vary the standard deviation? One way to do that would be

1. iterate over the indices
2. index into vectors of means and sds:

```{r}
sigma <- list(1, 5, 10)
str(sigma)
seq_along(mu) %>% 
  map(~rnorm(5, mu[[.]], sigma[[.]])) %>% 
  str()
```

But that obfuscates the intent of the code. Instead we could use map2() which iterates over two vectors in parallel:

```{r}
map2(mu, sigma, rnorm, n = 5) %>% str()
```

* the arguments that vary for each call come before the function
* arguments that are the same for every call come after the function

Like map( ), map2( ) is just a wrapper around a for loop:

```{r}
map2 <- function(x, y, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], y[[i]], ...)
  }
  out
}
```

You could also imagine map3( ), map4( ), map5( ), map6( ) etc, but that would get tedious quickly. 

Instead, purrr provides pmap( ) which takes a list of arguments. You might use that if you wanted to vary the mean, standard deviation, and number of samples:

```{r}
n <- list(1,3,5)
mu <- list(5, 10, -3)
sigma <- list(1, 5, 10)

args1 <- list(n, mu, sigma)
str(args1)

args1 %>% 
  pmap(rnorm) %>% 
  str()
```

If you don’t name the elements of list, pmap() will use positional matching when calling the function. That’s a little fragile, and makes the code harder to read, so it’s better to name the arguments:

```{r}
n <- list(1,3,5)
mu <- list(5, 10, -3)
sigma <- list(1, 5, 10)

args2 <- list(mean = mu, sd = sigma, n = n)
str(args2)
```

```{r}
args2 %>% 
  pmap(rnorm) %>% 
  str()
```

Since the arguments are all the same length, it makes sense to store them in a data frame:

```{r}
params <- tribble(
  ~mean, ~sd, ~n,
  5, 1, 1,
  10, 5, 3,
  -3, 10, 5
)
params
```

```{r}
params %>% 
  pmap(rnorm) %>% 
  str()
```

As soon as your code gets complicated, I think a data frame is a good approach because it ensures that each column has a name and is the same length as all the other columns.

## Invoking different functions

There’s one more step up in complexity - as well as varying the arguments to the function you might also vary the function itself:

```{r}
f <- c("runif", "rnorm", "rpois")
str(f)

param <- list(
  list(min = -1, max = 1),
  list(sd = 5),
  list(lambda = 10)
)
str(param)
```

To handle this case, you can use invoke_map():

```{r}
invoke_map(f, param, n = 5) %>% str()
```

1. The first argument is a list of functions or character vector of function names. 
2. The second argument is a list of lists giving the arguments that vary for each function. 
3. The subsequent arguments are passed on to every function.

And again, you can use tribble() to make creating these matching pairs a little easier:

```{r}
sim <- tribble(
  ~f, ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)

sim

sim <- sim %>% 
  mutate(result = invoke_map(f, params, n = 10))

str(sim$result)
```

# Walk

Walk is an alternative to map that you use when you want to call a function for its side effects, rather than for its return value. 

You typically do this because you want to render output to the screen or save files to disk - the important thing is the action, not the return value. Here’s a very simple example:

```{r}
x <- list(1, "a", 3)
str(x)
```

```{r}
x %>% 
  walk(print)
```

walk( ) is generally not that useful compared to walk2( ) or pwalk( ). 

For example, if you had a list of plots and a vector of file names, you could use pwalk( ) to save each file to the corresponding location on disk:

```{r}
library(ggplot2)

plots <- mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point())

paths <- stringr::str_c("../fig_output/", names(plots), ".pdf")

pwalk(list(paths, plots), ggsave, path = tempdir())
```

walk(), walk2() and pwalk() all invisibly return .x, the first argument. This makes them suitable for use in the middle of pipelines.

# Other patterns of for loops

Purrr provides a number of other functions that abstract over other types of for loops.

## Predicate functions

A number of functions work with predicate functions that return either a single TRUE or FALSE.

* keep( ) and discard( ) keep elements of the input where the predicate is TRUE or FALSE respectively:

```{r}
iris %>% 
  keep(is.factor)

iris %>% 
  keep(is.factor) %>% 
  str()

iris %>% 
  discard(is.factor)

iris %>% 
  discard(is.factor) %>% 
  str()
```

* some() and every() determine if the predicate is true for any or for all of the elements.

```{r}
x <- list(1:5, letters, list(10))
x
str(x)
```

```{r}
x %>% 
  some(is_character)
```

```{r}
x %>% 
  every(is_vector)
```

* detect( ) finds the first element where the predicate is true; detect_index( ) returns its position.

```{r}
x <- sample(10)
x
```

```{r}
x %>% 
  detect(~ . > 5)
```

```{r}
x %>% 
  detect_index(~ . > 5)
```

There is an anonymous function. purrr provides a convenient shortcut: a one-sided formula. And "." as a pronoun: it refers to the current list element (in the same way that i refereed to the current index in the for loop).

* head_while() and tail_while() take elements from the start or end of a vector while a predicate is true:

```{r}
x
x %>% 
  head_while(~ . > 5)
```

```{r}
x %>% 
  tail_while(~ . >5)
```

## Reduce and accumulate

Sometimes you have a complex list that you want to reduce to a simple list by repeatedly applying a function that reduces a pair to a singleton. This is useful if you want to apply a two-table dplyr verb to multiple tables. 

For example, you might have a list of data frames, and you want to reduce to a single data frame by joining the elements together:

```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)
dfs
str(dfs)
```

```{r}
dfs %>% 
  reduce(full_join)
```

Or maybe you have a list of vectors, and want to find the intersection:

```{r}
vs <- list(
  c(1, 3, 5, 6, 10),
  c(1, 2, 3, 7, 8, 10),
  c(1, 2, 3, 4, 8, 9, 10)
)
vs
str(vs)
```

```{r}
vs %>% 
  reduce(intersect)
```

The reduce function takes a “binary” function (i.e. a function with two primary inputs), and applies it repeatedly to a list until there is only a single element left.

Accumulate is similar but it keeps all the interim results. You could use it to implement a cumulative sum:

```{r}
x <- sample(10)
x

x %>% accumulate('+')
```

