---
title: "List Columns"
author: "Xiaochi"
date: "15/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(babynames)
```

```{r}
babynames
```

# Data Structures in R

```{r}
# collection
1
c(1, 2, 3.14)

# everything is a vector
is.vector(1)
is.vector("a")
is.vector(c(1, 2, 3.14))

# type of data
typeof(c(1L, 2L, 3L))
typeof(c(1, 2, 3.14))
typeof(c("a", "b", "c"))
typeof(c(TRUE, FALSE))

# attribute
x <- c(1L, 2L, 3L)
x
class(x) <- "Date"
x

levels(x) <- c("blue", "Brown", "Green")
class(x) <- "factor"
x

dim(x) <- c(3,1)
x

# list
y <- list(
  a = c(1, 2, 3.14),
  b = c("a", "b", "c"),
  c = c(TRUE, FALSE, FALSE)
)
y
typeof(y)
is.vector(y)
class(y) <- "data.frame"
rownames(y) <- c("1", "2", "3")
y
y$d <- list(p = 1:3, q = TRUE, r = 0L)
y
```

# Data frames, tibbles, and list columns

```{r}
data.frame(
  a = c(1, 2, 3.14),
  b = c("a", "b", "c"),
  c = c(TRUE, FALSE, FALSE)
)
```

```{r}
data.frame(
  a = c(1, 2, 3.14),
  b = c("a", "b", "c"),
  c = c(TRUE, FALSE, FALSE),
  d = list(p = 1:3, q = TRUE, r = 0L)
)
```

```{r}
z <- data.frame(
  a = c(1, 2, 3.14),
  b = c("a", "b", "c"),
  c = c(TRUE, FALSE, FALSE)
)
z
z$d <- list(p = 1:30, q = TRUE, r = 0L)
z
typeof(z)
class(z) <- c("tbl_df", "tbl", "data.frame")
typeof(z)
z
```

```{r}
library(tibble)
tibble(
  a = c(1, 2, 3.14),
  b = c("a", "b", "c"),
  c = c(TRUE, FALSE, FALSE),
  d = list(p = 1:3, q = TRUE, r = 0L)
)
```

```{r}
babynames
```

```{r}
as_tibble(babynames)
```

```{r}
as.data.frame((babynames))
```


# Single table verbs

* arrange: order rows by values of a column (low to high), use with desc() to order from high to low.

* filter: extract rows that meet logical criteria.

* select: extract columns by name.

* mutate: compute new column

* summarise: computate table of summaries. Use group_by() to compute groupwise summaries.

* nest and unnest

# Composing functions

pipes: x %>% f(y) becomes f(x, y)

```{r}
library(tidyverse)
babynames %>% 
  select(-prop) %>%
  filter(!is.na(n)) %>% 
  group_by(year, sex) %>% 
  summarise(N = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = N, color = sex))
```

# Map funtions

```{r}
y
```

```{r}
y %>% mutate(asq = sqrt(a))
```

* mutate works with tables
* sqrt works with structure stored in a column (double vector)

```{r error=TRUE}
y %>% mutate(asq = sqrt(d))
```

```{r}
y %>% mutate(dsq = map(d, sqrt))
```

* mutate works with tables
* map works with contents of column (list)
* sqrt works with contents of list (atomic vectors)

# Case Study

```{r}
everpresent <- babynames %>% 
  group_by(name, sex) %>% 
  summarise(years = n()) %>% 
  ungroup() %>% 
  filter(years == max(years))

babynames <- babynames %>% 
    semi_join(everpresent)
```

```{r}
babynames %>% 
  filter(name == "Joe", sex == "M") %>% 
  ggplot(aes(x = year, y = prop)) +
  geom_line() +
  geom_smooth(method = lm)
```

```{r}
babynames %>% 
  group_by(name, sex) %>% 
  nest() %>% 
  pluck("data") %>% 
  pluck(1)
```

```{r}
library(broom)
babymods <- babynames %>% 
  group_by(name, sex) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(prop~year, data = .x)),
         slope = map_dbl(model, ~pluck(coef(.x), "year")),
         r_squared = map_dbl(model, ~pluck(glance(.x), "r.squared"))
         )
babymods
```

Which names increased the most?

```{r}
babymods %>% 
  arrange(desc(slope)) %>% 
  head(5) %>% 
  unnest(data) %>% 
  ggplot(aes(x = year, y = prop)) +
  geom_line(aes(color = name))
```

which names were the least linear?

```{r}
babymods %>% 
  arrange(r_squared) %>% 
  head(5) %>% 
  unnest(data) %>% 
  ggplot(aes(x = year, y = prop)) +
  geom_line(aes(color = name))
```

# Tao of Tidy

