---
title: "Many models"
author: "Xiaochi"
date: "28/08/2019"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(modelr)
library(tidyverse)
library(gapminder)
```

three powerful ideas that help you to work with large numbers of models with ease:

1. Using many simple models to better understand complex datasets.

2. Using list-columns to store arbitrary data structures in a data frame. For example, this will allow you to have a column that contains linear models.

3. Using the broom package, by David Robinson, to turn models into tidy data. This is a powerful technique for working with large numbers of models because once you have tidy data, you can apply all of the techniques that you’ve learned about earlier in the book.

# gapminder

```{r}
gapminder
```

How does life expectancy (lifeExp) change over time (year) for each country (country)?

```{r}
gapminder %>% 
  count(country)
```

```{r}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
  geom_line(alpha = 1/3)
```

This is a small dataset: it only has ~1,700 observations and 3 variables. But it’s still hard to see what’s going on! Overall, it looks like life expectancy has been steadily improving. However, if you look closely, you might notice some countries that don’t follow this pattern. How can we make those countries easier to see?

One way is to use the same approach as in the last chapter: there’s a strong signal (overall linear growth) that makes it hard to see subtler trends.

We’ll tease these factors apart by fitting a model with a linear trend. The model captures steady growth over time, and the residuals will show what’s left.

You already know how to do that if we had a single country:

```{r}
nz <- filter(gapminder, country == "New Zealand")

nz
```

```{r}
nz %>% 
  ggplot(aes(year, lifeExp)) +
  geom_line() +
  ggtitle("Full data = ")

nz_mod <- lm(lifeExp ~ year, data = nz)

nz %>% 
  add_predictions(nz_mod) %>% 
  ggplot(aes(year, pred)) + 
  geom_line() +
  ggtitle("Linear trend + ")

nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) +
  geom_hline(yintercept = 0, colour = "white", size = 3) +
  geom_line() +
  ggtitle("Remaining pattern")
```

How can we easily fit that model to every country?

## Nested data

Extract out the common code with a function and repeat using a map function from purrr.

Instead of repeating an action for each variable, we want to repeat an action for each country, a subset of rows.

To do that, we need a new data structure: the nested data frame.

To create a nested data frame we start with a grouped data frame, and “nest” it:

```{r}
by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()
by_country
```

(I’m cheating a little by grouping on both continent and country. Given country, continent is fixed, so this doesn’t add any more groups, but it’s an easy way to carry an extra variable along for the ride.)

This creates a data frame that has one row per group (per country), and a rather unusual column: "data". 

"data" is a list of data frames (or tibbles, to be precise). This seems like a crazy idea: we have a data frame with a column that is a list of other data frames! 

if you pluck out a single element from the data column you’ll see that it contains all the data for that country (in this case, Afghanistan).

```{r}
by_country$data[[1]]
```

Note the difference between a standard grouped data frame and a nested data frame:

* in a grouped data frame, each row is an observation;
* in a nested data frame, each row is a group.

Another way to think about a nested dataset is we now have a meta-observation: a row that represents the complete time course for a country, rather than a single point in time.

## List-columns

Now that we have our nested data frame, we’re in a good position to fit some models. We have a model-fitting function:

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
```

And we want to apply it to every data frame. The data frames are in a list, so we can use purrr::map() to apply country_model to each element:

```{r}
models <- map(by_country$data, country_model)
length(models)
```

However, rather than leaving the list of models as a free-floating object, I think it’s better to store it as a column in the by_country data frame.

Storing related objects in columns is a key part of the value of data frames, and why I think list-columns are such a good idea.

In the course of working with these countries, we are going to have lots of lists where we have one element per country. So why not store them all together in one data frame?

In other words, instead of creating a new object in the global environment, we’re going to create a new variable in the by_country data frame. That’s a job for dplyr::mutate():

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}

by_country <- by_country %>% 
  mutate(model = map(data, country_model))

by_country
```

This has a big advantage: because all the related objects are stored together, you don’t need to manually keep them in sync when you filter or arrange. The semantics of the data frame takes care of that for you:

```{r}
by_country %>% 
  filter(continent == "Europe")
```

```{r}
by_country %>% 
  arrange(continent, country)
```

If your list of data frames and list of models were separate objects, you have to remember that whenever you re-order or subset one vector, you need to re-order or subset all the others in order to keep them in sync. If you forget, your code will continue to work, but it will give the wrong answer!

## Unnesting

Previously we computed the residuals of a single model with a single dataset. Now we have 142 data frames and 142 models. 

To compute the residuals, we need to call add_residuals() with each model-data pair:

```{r}
by_country <- by_country %>% 
  mutate(
    resids = map2(data, model, add_residuals)
  )
by_country
```

But how you can plot a list of data frames? 

Instead of struggling to answer that question, let’s turn the list of data frames back into a regular data frame.

* we used nest() to turn a regular data frame into an nested data frame
* we do the opposite with unnest():

```{r}
by_country$resids[[1]]
```

```{r}
by_country
```


```{r}
resids <- unnest(by_country, resids)
resids
```

Note that each regular column is repeated one for each row in the nested column.

Now we have regular data frame, we can plot the residuals:

```{r}
resids %>% 
  ggplot(aes(year, resid)) +
  geom_line(aes(group = country), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

Facetting by continent is particularly revealing:

```{r}
resids %>% 
  ggplot(aes(year, resid, group = country)) +
  geom_line(alpha = 1/3) +
  facet_wrap(~continent)
```

It looks like we’ve missed some mild patterns. There’s also something interesting going on in Africa: we see some very large residuals which suggests our model isn’t fitting so well there. 

We’ll explore that more in the next section, attacking it from a slightly different angle.

## Model quality

Instead of looking at the residuals from the model, we could look at some general measurements of model quality.

The broom package provides a general set of functions to turn models into tidy data.

Here we’ll use broom::glance() to extract some model quality metrics. If we apply it to a model, we get a data frame with a single row:

```{r}
nz_mod
```


```{r}
broom::glance(nz_mod)
```

```{r}
glance_df <- by_country %>% 
  mutate(glance = map(model, broom::glance))

glance_df$glance[[1]]
```


```{r}
by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)
```

This isn’t quite the output we want, because it still includes all the list columns.

This is default behaviour when unnest() works on single row data frames. To suppress these columns we use .drop = TRUE:

```{r}
by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)
```


```{r}
glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE)
glance
```

(Pay attention to the variables that aren’t printed: there’s a lot of useful stuff there.)

With this data frame in hand, we can start to look for models that don’t fit well:

```{r}
glance %>% 
  arrange(r.squared) %>% 
  select(country, continent, r.squared)
```

The worst models all appear to be in Africa. Let’s double check that with a plot. 

Here we have a relatively small number of observations and a discrete variable, so geom_jitter() is effective:

```{r}
glance %>% 
  ggplot(aes(continent, r.squared)) +
  geom_point()
```


```{r}
glance %>% 
  ggplot(aes(continent, r.squared)) +
  geom_jitter(width = 0.5)
```

We could pull out the countries with particularly bad R2 and plot the data:

```{r}
bad_fit <- filter(glance, r.squared < 0.25)
bad_fit
```


```{r}
gapminder %>% 
  # return all rows from gapminder where there are matching values in bad_fit
  # keep just columns from gapminder
  semi_join(bad_fit, by = "country") %>% 
  ggplot(aes(year, lifeExp, colour = country)) +
  geom_line()
```

# List-columns

List-columns are implicit in the definition of the data frame: a data frame is a named list of equal length vectors.

A list is a vector, so it’s always been legitimate to use a list as a column of a data frame.

However, base R doesn’t make it easy to create list-columns, and data.frame() treats a list as a list of columns:.

```{r}
x = list(1:3, 3:5)
x
```


```{r}
data.frame(x = list(1:3, 3:5))
```



You can prevent data.frame() from doing this with I(), but the result doesn’t print particularly well:

```{r}
data.frame(
  x = I(list(1:3, 3:5)),
  y = c("1, 2", "3, 4, 5")
)
```

Tibble alleviates this problem by being lazier (tibble() doesn’t modify its inputs) and by providing a better print method:

```{r}
tibble(
  x = list(1:3, 3:5),
  y = c("1, 2", "3, 4, 5")
)
```

```{r}
tribble(
  ~x, ~y,
  1:3, "1, 2",
  3:5, "3, 4, 5"
)
```

List-columns are often most useful as intermediate data structure. They’re hard to work with directly, because most R functions work with atomic vectors or data frames, but the advantage of keeping related items together in a data frame is worth a little hassle.

Generally there are three parts of an effective list-column pipeline:

1. You create the list-column using 
    * nest( )
    * summarise( ) + list( )
    * mutate( ) + a map function

2. You create other intermediate list-columns by transforming existing list columns with map(), map2() or pmap().

3. You simplify the list-column back down to a data frame or atomic vector.

# Creating list-columns

Typically, you won’t create list-columns with tibble().

Instead, you’ll create them from regular columns, using one of three methods:

1. With tidyr::nest( ) to convert a grouped data frame into a nested data frame where you have list-column of data frames.

2. With mutate( ) and vectorised functions that return a list.

3. With summarise() and summary functions that return multiple results.

Alternatively, you might create them from a named list, using tibble::enframe().

## With nesting

nest() creates a nested data frame, which is a data frame with a list-column of data frames. 

In a nested data frame:

* each row is a meta-observation
* the other columns give variables that define the observation (like country and continent above)
* the list-column of data frames gives the individual observations that make up the meta-observation

There are two ways to use nest():

1. When applied to a grouped data frame, nest() keeps the grouping columns as is, and bundles everything else into the list-column:

```{r}
gapminder %>% 
  group_by(country, continent) %>% 
  nest()
```

2. You can also use it on an ungrouped data frame, specifying which columns you want to nest:

```{r}
gapminder %>% 
  nest(year:gdpPercap)
```

## From vectorised functions

Some useful functions take an atomic vector and return a list.

For example, in strings you learned about stringr::str_split() which takes a character vector and returns a list of character vectors. If you use that inside mutate, you’ll get a list-column:

```{r}
df <- tribble(
  ~x1,
  "a,b,c",
  "d,e,f,g"
)
df
```

```{r}
stringr::str_split(df$x1, ",")
```


```{r}
df %>% 
  mutate(x2 = stringr::str_split(x1, ","))
```

unnest( ) knows how to handle these lists of vectors:

```{r}
df %>% 
  mutate(x2 = stringr::str_split(x1, ",")) %>% 
  unnest()
```

(If you find yourself using this pattern a lot, make sure to check out tidyr::separate_rows() which is a wrapper around this common pattern).

Another example of this pattern is using the map(), map2(), pmap() from purrr. For example, we could take the final example from Invoking different functions and rewrite it to use mutate():

```{r}
sim <- tribble(
  ~f, ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
sim
```

```{r}
invoke_map(sim$f, sim$params, n = 10)
# The first argument: a list of functions, or character vector of function names
# The second argument: a list of lists giving the arguments that vary for each function
# The subsequent arguments: are passed on to every function.
```

* The first argument: 

```{r}
sim %>% 
  mutate(sims = invoke_map(f, params, n = 10))
```

Note that technically sim isn’t homogeneous because it contains both double and integer vectors. However, this is unlikely to cause many problems since integers and doubles are both numeric vectors.

## From multivalued summaries

One restriction of summarise() is that it only works with summary functions that return a single value. 

That means that you can’t use it with functions like quantile( ) that return a vector of arbitrary length:

```{r}
quantile(mtcars$cyl)
```


You can however, wrap the result in a list! This obeys the contract of summarise(), because each summary is now a list (a vector) of length 1.


```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg)))
```

```{r}
probs <- c(0.01, 0.25, 0.5, 0.75, 0.99)
list(probs)
mtcars %>% 
  group_by(cyl) %>% 
  summarise(p = list(probs), q = list(quantile(mpg, probs))) %>% 
  unnest()
```

## From a named list

Data frames with list-columns provide a solution to a common problem: what do you do if you want to iterate over both the contents of a list and its elements?

it’s often easier to make a data frame: one column can contain the elements, and one column can contain the list. 

An easy way to create such a data frame from a list is tibble::enframe().

```{r}
x <- list(
  a = 1:5,
  b = 3:4,
  c = 5:6
)
x
```

```{r}
df <- enframe(x)
df
```

The advantage of this structure is that it generalises in a straightforward way - names are useful if you have character vector of metadata, but don’t help if you have other types of data, or multiple vectors.

Now if you want to iterate over names and values in parallel, you can use map2():

```{r}
df %>% 
  mutate(
    smry = map2_chr(name, value, ~ stringr::str_c(.x, ": ", .y[1]))
  )
```

# Simplifying list-columns

The technique you’ll use to collapse back down to a simpler structure depends on whether you want a single value per element, or multiple values:

1. If you want a single value, use mutate() with map_lgl(), map_int(), map_dbl(), and map_chr() to create an atomic vector.

2. If you want many values, use unnest() to convert list-columns back to regular columns, repeating the rows as many times as necessary.

## List to vector

If you can reduce your list column to an atomic vector then it will be a regular column.

For example, you can always summarise an object with its type and length, so this code will work regardless of what sort of list-column you have:

```{r}
df <- tribble(
  ~x,
  letters[1:5],
  1:3,
  runif(5)
)
df
```

```{r}
df %>% mutate(
  type = map_chr(x, typeof),
  length = map_int(x, length)
)
```

This is the same basic information that you get from the default tbl print method, but now you can use it for filtering. This is a useful technique if you have a heterogeneous list, and want to filter out the parts aren’t working for you.

Don’t forget about the map_*() shortcuts - you can use map_chr(x, "apple") to extract the string stored in apple for each element of x. This is useful for pulling apart nested lists into regular columns. Use the .null argument to provide a value to use if the element is missing (instead of returning NULL):

```{r}
df <- tribble(
  ~x,
  list(a = 1, b = 2),
  list(a = 2, c = 4)
)
df
```

```{r}
df %>% mutate(
  a = map_dbl(x, "a"),
  b = map_dbl(x, "b", .null = NA_real_)
)
```

## Unnesting

unnest() works by repeating the regular columns once for each element of the list-column.

For example, in the following very simple example we repeat the first row 4 times (because there the first element of y has length four), and the second row once:

```{r}
df <- tibble(
  x = 1:2,
  y = list(1:4, 1)
)
df
```

```{r}
df$y
```


```{r}
df %>% 
  unnest(y)
```

This means that you can’t simultaneously unnest two columns that contain different number of elements:

```{r}
df1 <- tribble(
  ~x, ~y, ~z,
  1, c("a", "b"), 1:2,
  2, "c", 3
)
df1
```

```{r}
df1$y
```

```{r}
df1$z
```

```{r}
# Ok, because y and z have the same number of elements in every row
df1 %>% 
  unnest(y, z)
```

```{r}
df2 <- tribble(
  ~x, ~y, ~z,
  1, "a", 1:2,
  2, c("b", "c"), 3
)
df2
```

```{r}
df2$y
```

```{r}
df2$z
```

```{r}
df2 %>% unnest(y, z)
```


The same principle applies when unnesting list-columns of data frames. You can unnest multiple list-cols as long as all the data frames in each row have the same number of rows.

# Making tidy data with broom

The broom package provides three general tools for turning models into tidy data frames:

1. broom::glance(model) returns a row for each model:
    * Each column gives a model summary: either a measure of model quality, or complexity, or a combination of the two.

2. broom::tidy(model) returns a row for each coefficient in the model:
    * Each column gives information about the estimate or its variability.
    
3. broom::augment(model, data) returns a row for each row in data
    * adding extra values like residuals, and influence statistics.
    

# Hadley

## Nested data (tidyr)

```{r}
gapminder <- gapminder %>% 
  mutate(year1950 = year - 1950)

gapminder
```

```{r}
by_country <- gapminder %>% 
  group_by(continent, country) %>% 
  nest()

by_country
```

```{r}
by_country$data[[1]]
```

```{r}
country_model <- function(df) {
  lm(lifeExp ~ year1950, data = df)
}

models <- by_country %>% 
  mutate(
    mod = map(data, country_model)
  )

models
```


## Functional programming (purrr)

Functional programming emphasises the actions

```{r}
mtcars
```

```{r}
map_dbl(mtcars, mean)
```

```{r}
map_dbl(mtcars, median)
```

```{r}
funs <- list(mean, median, sd)
funs
```

```{r}
funs %>% 
  map(~ mtcars %>% map_dbl(.x))
```

```{r}
models
```

```{r}
models %>% filter(continent == "Africa")
```

## Models to tidy data (broom)

What data can we extract from a model?

1. R square (glance, model statistics) 
2. Intercept, Slope (tidy, parameter statistics)
3. Resid (augment, observation statistics)

```{r}
models <- models %>% 
  mutate(
    tidy = map(mod, broom::tidy),
    glance = map(mod, broom::glance),
    rsq = map_dbl(glance, "r.squared"),
    augment = map(mod, broom::augment)
  )

models
```

```{r}
models %>% 
  arrange(desc(rsq)) %>% 
  select(country, continent, rsq)
```

```{r}
models %>% 
  # reorder:
  # treats its first argument as a categorical variable
  # reorders its levels based on the values of a second variable
  ggplot(aes(rsq, reorder(country, rsq))) +
  geom_point(aes(colour = continent))
```

```{r}
models %>% 
  unnest(data)
```

```{r}
models %>% 
  unnest(tidy)
```

```{r}
models %>% 
  unnest(tidy) %>% 
  select(continent, country, term, estimate, rsq) %>% 
  pivot_wider(names_from = term, values_from = estimate) %>% 
  ggplot(aes(`(Intercept)`, year1950)) +
  geom_point(aes(colour = continent, size = rsq)) +
  geom_smooth(se = FALSE) +
  xlab("Life Expectancy (1950)") +
  ylab("Yearly improveent") +
  scale_size_area()
```


```{r}
models %>% 
  unnest(glance)
```

```{r}
models %>% 
  unnest(augment)
```

```{r}
models %>% 
  unnest(augment) %>% 
  ggplot(aes(year1950, .resid)) +
  geom_line(aes(group = country), alpha = 1 / 3) +
  geom_smooth(se = FALSE) +
  geom_hline(yintercept = 0, colour = "white") +
  facet_wrap(~continent)
```

1. Store related objects in list-columns.
2. Learn FP so you can focus on verbs, not objects.
3. Use broom to convert models to tidy data.