---
title: "Tibbles"
author: "Xiaochi"
date: "09/08/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(tidyverse)
```

# Creating tibbles

```{r}
iris
```

```{r}
as_tibble(iris)
```

```{r}
tibble(
  x=1:5,
  y=1,
  z=x^2+y
)
```

tibble() does much less: it never changes the type of the inputs (e.g. it never converts strings to factors!), it never changes the names of variables, and it never creates row names.

Another way to create a tibble is with tribble(), short for transposed tibble. tribble() is customised for data entry in code: column headings are defined by formulas (i.e. they start with ~), and entries are separated by commas. This makes it possible to lay out small amounts of data in easy to read form.

```{r}
tribble(
  ~x,~y,~z,
  "a",2,3.6,
  "b",1,8.5
)
```

# Tibbles vs. data.frame

There are two main differences in the usage of a tibble vs. a classic data.frame: printing and subsetting.

## Printing

Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen. This makes it much easier to work with large data. In addition to its name, each column reports its type, a nice feature borrowed from str():

```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

Tibbles are designed so that you don’t accidentally overwhelm your console when you print large data frames. But sometimes you need more output than the default display. There are a few options that can help.

First, you can explicitly print() the data frame and control the number of rows (n) and the width of the display. width = Inf will display all columns:

```{r}
nycflights13::flights %>% 
  print(n=10,width=Inf)
```

You can also control the default print behaviour by setting options:

* options(tibble.print_max = n, tibble.print_min = m): if more than n rows, print only m rows. Use options(tibble.print_min = Inf) to always show all rows.

* Use options(tibble.width = Inf) to always print all columns, regardless of the width of the screen.

A final option is to use RStudio’s built-in data viewer to get a scrollable view of the complete dataset. This is also often useful at the end of a long chain of manipulations.

## Subsetting

```{r}
df <- tibble(
  x=runif(5),
  y=rnorm(5)
)

# Extract by name
df$x
df[["x"]]

# Extract by position
df[[1]]
```

To use these in a pipe, you’ll need to use the special placeholder .:

```{r}
df %>% .$x

df %>% .[["x"]]
```

Compared to a data.frame, tibbles are more strict: they never do partial matching, and they will generate a warning if the column you are trying to access does not exist.

# Interacting with older code

Compared to a data.frame, tibbles are more strict: they never do partial matching, and they will generate a warning if the column you are trying to access does not exist.

The main reason that some older functions don’t work with tibble is the [ function. We don’t use [ much in this book because dplyr::filter() and dplyr::select() allow you to solve the same problems with clearer code (but you will learn a little about it in vector subsetting). With base R data frames, [ sometimes returns a data frame, and sometimes returns a vector. With tibbles, [ always returns another tibble.




