---
title: "Vectors"
author: "Xiaochi"
date: "19/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(tidyverse)
```

Vectors are particularly important as most of the functions you will write will work with vectors. It is possible to write functions that work with tibbles (like ggplot2, dplyr, and tidyr), but the tools you need to write such functions are currently idiosyncratic and immature. I am working on a better approach, https://github.com/hadley/lazyeval, but it will not be ready in time for the publication of the book. Even when complete, you’ll still need to understand vectors, it’ll just make it easier to write a user-friendly layer on top.

# Vector basics

There are two types of vectors:

1. Atomic vectors
    + logical
    + integer
    + double
    + character
    + complex
    + raw

2. Lists

atomic vectors are homogeneous, while lists can be heterogeneous.

There’s one other related object: NULL. 

* NULL is often used to represent the absence of a vector
* as opposed to NA which is used to represent the absence of a value in a vector). 

NULL typically behaves like a vector of length 0.

Every vector has two key properties:

1. type,typeof( ).

```{r}
typeof(letters)
typeof(1:10)
```

2. length, length( ).

```{r}
x <- list("a", "b", 1:10)
length(x)
```

Vectors can also contain arbitrary additional metadata in the form of attributes. These attributes are used to create augmented vectors which build on additional behaviour. There are three important types of augmented vector:

* Factors -- integer vectors.
* Dates and date-times -- numeric vectors.
* Data frames and tibbles -- lists.

# Import types of atomic vector

The four most important types of atomic vector are 

* logical
* integer
* double
* character

## Logical

three possible values: 

* FALSE
* TRUE
* NA. 

Logical vectors are usually constructed with comparison operators, as described in comparisons. 

```{r}
1:10 %% 3 == 0
```

You can also create them by hand with c( ):

```{r}
c(TRUE, TRUE, FALSE, NA)
```

## Numeric

Integer and double vectors are known collectively as numeric vectors. 

In R, numbers are doubles by default. 

To make an integer, place an L after the number:

```{r}
typeof(1)
typeof(1L)
1.5L
```

The distinction between integers and doubles is not usually important, but there are two important differences that you should be aware of:

1. Doubles are approximations. Doubles represent floating point numbers that can not always be precisely represented with a fixed amount of memory. This means that you should consider all doubles to be approximations. For example, what is square of the square root of two?

```{r}
x <- sqrt(2) ^ 2
x
typeof(x)
x-2
```

This behaviour is common when working with floating point numbers: most calculations include some approximation error. 

Instead of comparing floating point numbers using ==, you should use dplyr::near( ) which allows for some numerical tolerance.

2. Integers have one special value: NA, while doubles have four: 

* NA
* NaN 
* Inf
* -Inf

```{r}
c(-1, 0, 1) / 0
```

Avoid using == to check for these other special values. 

Instead use the helper functions 

* is.finite( )
* is.infinite( )
* is.nan( )

## Character

each element of a character vector is a string
a string can contain an arbitrary amount of data.

R uses a global string pool. This means that each unique string is only stored in memory once, and every use of the string points to that representation. 

This reduces the amount of memory needed by duplicated strings. You can see this behaviour in practice with pryr::object_size():

```{r}
x <- "This is a reasonably long string."
pryr::object_size(x)

y <- rep(x, 1000)
pryr::object_size(y)
```

y doesn’t take up 1,000x as much memory as x, because each element of y is just a pointer to that same string. A pointer is 8 bytes, so 1000 pointers to a 136 B string is 8 * 1000 + 136 = 8.13 kB.

## Missing values

each type of atomic vector has its own missing value:

```{r}
NA
typeof(NA)

NA_integer_
typeof(NA_integer_)

NA_real_
typeof(NA_real_)

NA_character_
typeof(NA_character_)
```

# Using atomic vectors

some of the important tools for working with them. These include:

1. How to convert from one type to another, and when that happens automatically.

2. How to tell if an object is a specific type of vector.

3. What happens when you work with vectors of different lengths.

4. How to name the elements of a vector.

5. How to pull out elements of interest.

## Coercion

There are two ways to convert, or coerce, one type of vector to another:

1. Explicit coercion
    + as.logical( )
    + as.integer( )
    + as.double( )
    + as.character( ) 

2. Implicit coercion
    + in a specific context that expects a certain type of vector
    + use a logical vector with a numeric summary function
    + use a double vector where an integer vector is expected

For example, using a logical vector in a numeric context. In this case TRUE is converted to 1 and FALSE converted to 0. That means the sum of a logical vector is the number of trues, and the mean of a logical vector is the proportion of trues:

```{r}
x <- sample(20, 100, replace = TRUE)
x
y <- x > 10
y

# how many are greater than 10?
sum(y) 
# what proportion are greater than 10?
mean(y)
```

It’s also important to understand what happens when you try and create a vector containing multiple types with c(): the most complex type always wins.

```{r}
typeof(c(TRUE, 1L))
typeof(c(1L,1.5))
typeof(c(1.5,"a"))
```

An atomic vector can not have a mix of different types because the type is a property of the complete vector, not the individual elements. 

If you need to mix multiple types in the same vector, you should use a list, which you’ll learn about shortly.

## Test functions

Sometimes you want to do different things based on the type of vector. 

1.  typeof( ). 
2.  is_*
    + is_logical( )
    + is_integer( )
    + is_double( )
    + is_numeric( )
    + is_character( )
    + is_atomic( )
    + is_list( )
    + is_vector( )

Each predicate also comes with a “scalar” version, like is_scalar_atomic(), which checks that the length is 1. This is useful, for example, if you want to check that an argument to your function is a single logical value.

## Scalars and recycling rules

As well as implicitly coercing the types of vectors to be compatible, R will also implicitly coerce the length of vectors. 

This is called vector recycling, because the shorter vector is repeated, or recycled, to the same length as the longer vector.

This is generally most useful when you are mixing vectors and scalar, a sigle number.

R doesn’t actually have scalar, instead, a single number is a vector of length 1. 

Because there are no scalars, most built-in functions are vectorised, meaning that they will operate on a vector of numbers. That’s why, for example, this code works:

```{r}
x <- sample(10)
x
x + 100
runif(10) > 0.5
```

In R, basic mathematical operations work with vectors. That means that you should never need to perform explicit iteration when performing simple mathematical computations.

It’s intuitive what should happen if you add two vectors of the same length, or a vector and a scalar, but what happens if you add two vectors of different lengths?

```{r}
# the longer vector
1:10

# the shorter vector
1:2

1:10 + 1:2
```

Here, R will expand the shortest vector to the same length as the longest, so called recycling. 

This is silent except when the length of the longer is not an integer multiple of the length of the shorter:

```{r}
1:10

1:3

1:10 + 1:3
```

While vector recycling can be used to create very succinct, clever code, it can also silently conceal problems. For this reason, the vectorised functions in tidyverse will throw errors when you recycle anything other than a scalar. If you do want to recycle, you’ll need to do it yourself with rep():

```{r}
tibble(x = 1:4, y = rep(1:2, 2))
```

```{r}
tibble(x = 1:4, y = rep(1:2, each = 2))
```

## Naming vectors

All types of vectors can be named. 

* You can name them during creation with c():

```{r}
c(x = 1, y = 2, z = 4)
```

* Or after the fact with purrr::set_names():

```{r}
set_names(1:3, c("a", "b", "c"))
```

Named vectors are most useful for subsetting, described next.

## Subsetting

So far we’ve used dplyr::filter() to filter the rows in a tibble. filter() only works with tibble, so we’ll need new tool for vectors: [. 

[ is the subsetting function, and is called like x[a]. There are four types of things that you can subset a vector with:

1. Subsetting with a numeric vector:
    + The integers must either be all positive, all negative, or zero.
    + Subseting with positive integers keeps the elements at those positions;
    + By repeating a position, you can actually make a longer output than input.
    + Negative values drop the elements at the specified positions.
    + It's an error to mix positive and negative values.
    + subseting with zero returns no value.

```{r}
x <- c("one", "two", "three", "four", "five")
x
x[c(3, 2, 5)]
x[c(1, 1, 5, 5, 5, 2)]
x[c(-1, -3, -5)]
x[0]
```


2. Subsetting with a logical vector:
    + keeps all values corresponding to a TRUE value.

```{r}
x <- c(10, 3, NA, 5, 8, 1, NA)
x

# All non-missing values of x
x[!is.na(x)]

# All even (or missing!) values of x
x[x %% 2 == 0]
```

3. Subsetting with a character vector:

```{r}
x <- c(abc = 1, def = 2, xyz = 5)
x
x[c("xyz", "def")]
typeof(x)
```

4. Subsetting with nothing, x[]:
    + returns the complete x. 
    + select all the rows or all the columns
    + x[1, ] selects the first row and all the columns
    + x[, -1] selects all rows and all columns except the first

There is an important variation of [ called [[. 

 * [[ only ever extracts a single element
 * always drops names. 
 
It’s a good idea to use [[]] whenever you want to make it clear that you’re extracting a single item, as in a for loop. The distinction between [ and [[ is most important for lists, as we’ll see shortly.

# Recursive vectors (lists)

Lists can contain other lists. This makes them suitable for representing hierarchical or tree-like structures. 

You create a list with list():

```{r}
x <- list(1,2,3)
x
```

A very useful tool for working with lists is str() because it focusses on the structure, not the contents.

```{r}
str(x)
```

```{r}
x_named <- list(a = 1, b = 2, c = 3)
x_named
str(x_named)
```

Unlike atomic vectors, list() can contain a mix of objects:

```{r}
y <- list("a", 1L, 1.5, TRUE)
y
str(y)
```

Lists can even contain other lists!

```{r}
z <- list(list(1, 2), list(3, 4))
z
str(z)
```

## Visualising lists

```{r}
x1 <- list(c(1, 2), c(3, 4))
x1
str(x1)
```


```{r}
x2 <- list(list(1, 2), list(3, 4))
x2
str(x2)
```


```{r}
x3 <- list(1, list(2, list(3)))
x3
str(x3)
```

## Subsetting

There are three ways to subset a list, which I’ll illustrate with a list named a:

```{r}
a <- list(a = 1:3, b = "a string", c = pi, d = list(-1, -5))
a
str(a)
```

* [ extracts a sub-list. The result will always be a list.

```{r}
a[1:2]
str(a[1:2])
```

```{r}
a[4]
str(a[4])
```

Like with vectors, you can subset with a logical, integer, or character vector.

* [[ extracts a single component from a list. It removes a level of hierarchy from the list.

```{r}
a[[1]]
str(a[[1]])
```

```{r}
a[[2]]
str(a[[2]])
```

```{r}
a[[3]]
str(a[[3]])
```

```{r}
a[[4]]
str(a[[4]])
```

* $ is a shorthand for extracting named elements of a list. It works similarly to [[ except that you don’t need to use quotes.

```{r}
a$a
a[["a"]]
```

```{r}
a$d
a[[4]]
```

* The distinction between [ and [[ is really important for lists
    + [ returns a new, smaller list
    + [[ drills down into the list

* If this pepper shaker is your list x: 
    + x[1] is a pepper shaker containing a single pepper packet. 
    + x[2] would look the same, but would contain the second packet. 
    + x[1:2] would be a pepper shaker containing two pepper packets.
    + x[[1]] is a single peper packet.
    + x[[1]][[1]] is the content of the pepper pacage.

# Attributes

Any vector can contain arbitrary additional metadata through its attributes. 

You can think of attributes as named list of vectors that can be attached to any object. 

You can get and set individual attribute values with attr() or see them all at once with attributes().

```{r}
x <- 1:10
x
attr(x, "greeting")
(attr(x, "greeting") <- "Hi!")
(attr(x, "farewell") <- "Bye!")

attributes(x)
```

# Augmented vectors

Atomic vectors and lists are the building blocks for other important vector types like factors and dates. I call these augmented vectors, because they are vectors with additional attributes, including class. 

Because augmented vectors have a class, they behave differently to the atomic vector on which they are built. In this book, we make use of four important augmented vectors:

* Factors
* Dates
* Date-times
* Tibbles

## Factors

* Factors are designed to represent categorical data that can take a fixed set of possible values. 
    + Factors are built on top of integers
    + have a levels attribute

```{r}
x <- factor(c("ab", "cd", "ab"), levels = c("ab", "cd", "ef"))
x
typeof(x)
attributes(x)
```

## Date and date-times

Dates in R are numeric vectors that represent the number of days since 1 January 1970.

```{r}
x <- as.Date("1971-01-01")

x
typeof(x)
unclass(x)
typeof(x)
attributes(x)
```

Date-times are numeric vectors with class POSIXct that represent the number of seconds since 1 January 1970. (In case you were wondering, “POSIXct” stands for “Portable Operating System Interface”, calendar time.)

```{r}
x <- lubridate::ymd_hm("1970-01-01 01:00")

x
unclass(x)
typeof(x)
attributes(x)
```

The tzone attribute is optional. It controls how the time is printed, not what absolute time it refers to.

```{r}
attr(x, "tzone") <- "US/Pacific"
x

attr(x, "tzone") <- "US/Eastern"
x
```

There is another type of date-times called POSIXlt. These are built on top of named lists:

```{r}
y <- as.POSIXlt(x)
y
typeof(y)
attributes(y)
```

## Tibbles

* Tibbles are augmented lists
    + they have class “tbl_df” + “tbl” + “data.frame”
    + names (column) attribute
    + row.names attribute

```{r}
tb <- tibble::tibble(x = 1:5, y = 5:1)
tb
typeof(tb)
attributes(tb)
```

The difference between a tibble and a list is that all the elements of a data frame must be vectors with the same length. All functions that work with tibbles enforce this constraint.

Traditional data.frames have a very similar structure:

```{r}
df <- data.frame(x = 1:5, y = 5:1)
df
typeof(df)
attributes(df)
```

The main difference is the class. The class of tibble includes “data.frame” which means tibbles inherit the regular data frame behaviour by default.

