---
title: "ANCOVA"
author: "Xiaochi"
date: "20/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(rlang)
library(tidyverse)
library(R.oo)
```

```{r}
ex1 <- expr(x + y)
ex2 <- expr(!!ex1 + z)
ex2 <- expr(ex1 + z)

ex3 <- expr(1 / !!ex1)
ex3 <- expr(1 / ex1)
```

```{r}
f1 <- function(z) expr(z)
f1(x + y)
```

```{r}
f2 <- function(z) enexpr(z)
f2(x +y)
```

```{r}
df <- data.frame(x = 1:5, y = 5:1)
df
```

```{r}
filter(df, abs(x) > 1e-3)
filter(df, abs(y) > 1e-3)

my_filter <- function(df, var) {
  var <- enquo(var)
  filter(df, abs(!!var) > 1e-3)
}
my_filter(df, x)
```

# Example 1: mtcars data

```{r}
input <- mtcars %>% 
  select(am, hp, mpg) %>% 
  mutate(am = as_factor(am))

summary(input)
str(input)
```

am (transmission type) - Cateforical Variable
hp (horse power) - Continous predictor variable
mpg (Miles per Gallon) - Response Variable

* To fit the model, we can use lm or aov
* aov: Fit an analysis of variance model by a call to lm for each stratum.

## Model1: Interaction between hp and am

```{r}
summary(lm(mpg ~ hp * am - 1,data = input))
```

```{r}
fit1 <- aov(mpg ~ hp * am,data = input)
summary(fit1)
```

* Both hp and am has significant effect on mpg, as p value in both cases is less than 0.05

* Interaction between these two variables is not significant as the p-value is more than 0.05

## Model2: No interaction between hp and am

```{r}
fit2 <- aov(mpg ~ hp + am,data = input)
summary(fit2)
```

* both hp and am has significant effect on mpg, as the p value in both cases is less than 0.05

## Comparing Two Models

Objective: to conclude if the interaction of the variables is truly in-significant

```{r}
anova(fit1, fit2)
```

* As the p-value is greater than 0.05, we conclude that the interaction between hp and am is not significant

* Therefore, mpg will depend in a similar manner on hp in both auto and manual transmission mode.

# Example 2: Swiss Fertility data

```{r}
input <- swiss %>% 
  select(Fertility, Agriculture, Catholic) %>% 
  mutate(Catholic = as.numeric(cut(Catholic, 2)))

input
```

## Model1: Interaction between Agriculture and Catholic

```{r}
fit1 <- aov(Fertility ~ Agriculture * Catholic,
            data = input)
summary(fit1)
```

## Model2: No Interaction between Agriculture and Catholic

```{r}
fit2 <- aov(Fertility ~ Agriculture + Catholic,
            data = input)
summary(fit2)
```

## Comparing Two Models

```{r}
anova(fit1, fit2)
```

* As the p-value is greater than 0.05, we conclude that the interaction between Agriculture and Catholic is not significant.

# Comparing Regression Lines

Objective: comparing the typical lifetime (hours) of two tool types (A and B).


```{r}
tools <- read.table(
  "http://statacumen.com/teach/ADA2/ADA2_notes_Ch07_tools.dat" , 
  header = TRUE)

summary(tools)
```

```{r}
ggplot(tools, aes(x = type, y = lifetime)) +
  geom_hline(
    aes(yintercept = mean(lifetime)), 
    colour = "black", 
    linetype = "dashed", 
    size = 0.3, 
    alpha = 0.5) +
  geom_boxplot(size = 0.75, alpha = 0.5) +
  geom_point(position = position_jitter(w = 0.05, h = 0), alpha = 0.5) +
  stat_summary(fun.y = mean, geom = "point", shape = 18, size = 6, colour="red", alpha = 0.8) +
  stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = .2, colour="red", alpha = 0.8) +
  labs(title = "Tool type lifetime") + 
  ylab("lifetime (hours)") +
  coord_flip()
#
```

A two sample t-test comparing mean lifetimes of tool types indicates a difference between means.

```{r}
t.summary <- t.test(lifetime ~ type, data = tools)
t.summary
```

```{r}
toolsfake <- read.table(
  "http://statacumen.com/teach/ADA2/ADA2_notes_Ch07_toolsfake.dat" , 
  header = TRUE)


ggplot(toolsfake, 
       aes(x = speed, y = hours, colour = type, shape = type)) +
  geom_point(size=4) +
  scale_shape_manual(values=charToInt(sort(unique(toolsfake$type)))) +
  labs(title="Fake tools data, hours by speed with categorical type") 
```

For the tool lifetime problem, you should compare groups (tools) after ad-
justing the lifetimes to account for the influence of a measurement variable, speed. The appropriate statistical technique for handling this problem is called analysis of covariance (ANCOVA).

## ANCOVA

```{r}
Input = ("
Species   Temp   Pulse
 ex       20.8   67.9
 ex       20.8   65.1
 ex       24     77.3
 ex       24     78.7
 ex       24     79.4
 ex       24     80.4
 ex       26.2   85.8
 ex       26.2   86.6
 ex       26.2   87.5
 ex       26.2   89.1
 ex       28.4   98.6
 ex       29    100.8
 ex       30.4   99.3
 ex       30.4  101.7
 niv      17.2   44.3
 niv      18.3   47.2
 niv      18.3   47.6
 niv      18.3   49.6
 niv      18.9   50.3
 niv      18.9   51.8
 niv      20.4   60
 niv      21     58.5
 niv      21     58.9
 niv      22.1   60.7
 niv      23.5   69.8
 niv      24.2   70.9
 niv      25.9   76.2
 niv      26.5   76.1
 niv      26.5   77
 niv      26.5   77.7
 niv      28.6   84.7
")

Data <-  read.table(textConnection(Input),header=TRUE)
Data <- as_tibble(Data)
```

```{r}
Data %>% 
  ggplot(aes(x = Temp, y = Pulse, color = Species)) +
  geom_line()
```

```{r}
library(car)
```


```{r}
model1 <- lm(Pulse ~ Temp + Species + Temp:Species, data = Data)
Anova(model1, type = "II")
```

Interaction is not significant

```{r}
model2 <- lm(Pulse ~ Temp + Species, data = Data)
Anova(model2, type = "II")
```

```{r}
summary(model2)
```

```{r}
I.nought = -7.21091
I1 = I.nought + 0
I2 = I.nought + -10.06529
B  = 3.60275
```

```{r}
Data %>% 
  ggplot(aes(Temp, Pulse, colour = Species)) +
  geom_point() +
  geom_abline(slope = B, intercept = I1) +
  geom_abline(slope = B, intercept = I2)
```

```{r}
summary(model2)
```


```{r}
gator
```




