---
title: "Do Test"
author: "Xiaochi"
date: "09/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(reprex)
library(ggpubr)
library(tidyverse)
library(rstatix)
options(help_type = "html")
```

## normality test

```{r}
# Store the data in the variable my_data
my_data <- ToothGrowth
```

```{r}
set.seed(1234)
sample_n(my_data, 10)
```

```{r}
ggdensity(my_data$len, 
          main = "Density plot of tooth length",
          xlab = "Tooth length")
```

```{r}
ggqqplot(my_data$len)
```

```{r}
library("car")
qqPlot(my_data$len)
```

```{r}
shapiro.test(my_data$len)
```

From the output, the p-value > 0.05 implying that the distribution of the data are not significantly different from normal distribution. 
In other words, we can assume the normality.


## one-sample t-test

```{r}
set.seed(1234)
my_data <- data.frame(
  name = paste0(rep("M_", 10), 1:10),
  weight = round(rnorm(10, 20, 2), 1)
)
```

```{r}
summary(my_data$weight)
```

```{r}
ggboxplot(my_data$weight, 
          ylab = "Weight (g)", xlab = FALSE,
          ggtheme = theme_minimal())
```

```{r}
shapiro.test(my_data$weight)
```

From the output, the p-value is greater than the significance level 0.05 implying that the distribution of the data are not significantly different from normal distribtion. 
In other words, we can assume the normality.

```{r}
ggqqplot(my_data$weight, 
         ylab = "Men's weight",
         ggtheme = theme_minimal())
```

From the normality plots, we conclude that the data may come from normal distributions.

```{r}
# One-sample t-test
res <- t.test(my_data$weight, mu = 25, alternative = "two.sided")
```

The p-value of the test is 7.95310^{-6}, which is less than the significance level alpha = 0.05. 
We can conclude that the mean weight of the mice is significantly different from 25g with a p-value = 7.95310^{-6}.

```{r}
t.test(my_data$weight, mu = 25, alternative = "less")
```

```{r}
t.test(my_data$weight, mu = 25, alternative = "greater")
```

```{r}
res$p.value
res$estimate
res$conf.int
```

## one-sample Wilcoxon signed rank test

```{r}
set.seed(1234)
my_data <- data.frame(
  name = paste0(rep("M_", 10), 1:10),
  weight = round(rnorm(10, 20, 2), 1)
)
```

```{r}
# Statistical summaries of weight
summary(my_data$weight)
```

```{r}
ggboxplot(my_data$weight, 
          ylab = "Weight (g)", xlab = FALSE,
          ggtheme = theme_minimal())
```

```{r}
# One-sample wilcoxon test
res <- wilcox.test(my_data$weight, mu = 25)
# Printing the results
res 
```

```{r}
# print only the p-value
res$p.value
```

The p-value of the test is 0.005793, which is less than the significance level alpha = 0.05. 
We can reject the null hypothesis and conclude that the average weight of the mice is significantly different from 25g with a p-value = 0.005793.

```{r}
wilcox.test(my_data$weight, mu = 25, alternative = "less")
```

```{r}
wilcox.test(my_data$weight, mu = 25, alternative = "greater")
```

## unpaired two-sample t-test

```{r}
# Data in two numeric vectors
women_weight <- c(38.9, 61.2, 73.3, 21.8, 63.4, 64.6, 48.4, 48.8, 48.5)
men_weight <- c(67.8, 60, 63.4, 76, 89.4, 73.3, 67.3, 61.3, 62.4) 
# Create a data frame
my_data <- data.frame( 
                group = rep(c("Woman", "Man"), each = 9),
                weight = c(women_weight,  men_weight)
                )
```

```{r}
my_data %>% 
  group_by(group) %>% 
  summarise(
    count = n(),
    mean = mean(weight, na.rm = TRUE),
    sd = sd(weight, na.rm = TRUE)
  )
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", 
          palette = c("#00AFBB", "#E7B800"),
          ylab = "Weight", xlab = "Groups")
```

```{r}
# Shapiro-Wilk normality test for Men's weights
with(my_data, shapiro.test(weight[group == "Man"]))
# Shapiro-Wilk normality test for Women's weights
with(my_data, shapiro.test(weight[group == "Woman"]))
```

From the output, the two p-values are greater than the significance level 0.05 implying that the distribution of the data are not significantly different from the normal distribution. 
In other words, we can assume the normality.

```{r}
var.test(weight ~ group, data = my_data)
```

The p-value of F-test is p = 0.1713596. 
It’s greater than the significance level alpha = 0.05. 
In conclusion, there is no significant difference between the variances of the two sets of data. 
Therefore, we can use the classic t-test witch assume equality of the two variances.

```{r}
# Compute t-test
res <- t.test(women_weight, men_weight, var.equal = TRUE)
res
```

```{r}
# Compute t-test
res <- t.test(weight ~ group, data = my_data, var.equal = TRUE)
res
```

The p-value of the test is 0.01327, which is less than the significance level alpha = 0.05. We can conclude that men’s average weight is significantly different from women’s average weight with a p-value = 0.01327.

```{r}
t.test(weight ~ group, data = my_data,
        var.equal = TRUE, alternative = "less")
```

```{r}
t.test(weight ~ group, data = my_data,
        var.equal = TRUE, alternative = "greater")
```

```{r}
# printing the p-value
res$p.value
```

```{r}
# printing the mean
res$estimate
```

```{r}
# printing the confidence interval
res$conf.int
```

## unpaired two-sample Wilcoxon test

```{r}
# Data in two numeric vectors
women_weight <- c(38.9, 61.2, 73.3, 21.8, 63.4, 64.6, 48.4, 48.8, 48.5)
men_weight <- c(67.8, 60, 63.4, 76, 89.4, 73.3, 67.3, 61.3, 62.4) 
# Create a data frame
my_data <- data.frame( 
                group = rep(c("Woman", "Man"), each = 9),
                weight = c(women_weight,  men_weight)
                )
```

```{r}
my_data %>% 
  group_by(group) %>% 
  summarise(
    count = n(),
    median = median(weight, na.rm = TRUE),
    IQR = IQR(weight, na.rm = TRUE)
  )
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800"),
          ylab = "Weight", xlab = "Groups")
```

```{r}
res <- wilcox.test(women_weight, men_weight)
res
```

```{r}
res <- wilcox.test(weight ~ group, data = my_data,
                   exact = FALSE)
res
```

```{r}
# Print the p-value only
res$p.value
```

The p-value of the test is 0.02712, which is less
than the significance level alpha = 0.05. 
We can conclude that men’s median weight is significantly different from women’s median weight with a p-value = 0.02712.

```{r}
wilcox.test(weight ~ group, data = my_data, 
        exact = FALSE, alternative = "less")
```

```{r}
wilcox.test(weight ~ group, data = my_data,
        exact = FALSE, alternative = "greater")
```

## paired sample t-test

```{r}
# Weight of the mice before treatment
before <-c(200.1, 190.9, 192.7, 213, 241.4, 196.9, 172.2, 185.5, 205.2, 193.7)
# Weight of the mice after treatment
after <-c(392.9, 393.2, 345.1, 393, 434, 427.9, 422, 383.9, 392.3, 352.2)
# Create a data frame
my_data <- data.frame( 
                group = rep(c("before", "after"), each = 10),
                weight = c(before,  after)
                )
```

```{r}
my_data %>% 
  group_by(group) %>% 
  summarise(
    count = n(),
    mean = mean(weight, na.rm = TRUE),
    sd = sd(weight, na.rm = TRUE)
  )
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800"),
          order = c("before", "after"),
          ylab = "Weight", xlab = "Groups")
```

```{r}
library(PairedData)
```

```{r}
# Subset weight data before treatment
before <- subset(my_data, group == "before", weight, drop = TRUE)
# subset weight data after treatment
after <- subset(my_data, group == "after", weight, drop = TRUE)

pd <- paired(before, after)
plot(pd, type = "profile") + theme_bw()
```

```{r}
# compute the difference
d <- with(my_data, weight[group == "before"] - weight[group == "after"])
# Shapiro-Wilk normality test for the differences
shapiro.test(d) # => p-value = 0.6141
```

From the output, the p-value is greater than the significance level 0.05 implying that the distribution of the differences (d) are not significantly different from normal distribution. 
In other words, we can assume the normality.

```{r}
# Compute t-test
res <- t.test(before, after, paired = TRUE)
res
```

The p-value of the test is 6.210^{-9}, which is less than the significance level alpha = 0.05. We can then reject null hypothesis and conclude that the average weight of the mice before treatment is significantly different from the average weight after treatment with a p-value = 6.210^{-9}.

```{r}
# Compute t-test
res <- t.test(weight ~ group, data = my_data, paired = TRUE)
res
```

The p-value of the test is 6.210^{-9}, which is less than the significance level alpha = 0.05. We can then reject null hypothesis and conclude that the average weight of the mice before treatment is significantly different from the average weight after treatment with a p-value = 6.210^{-9}.

```{r}
t.test(weight ~ group, 
       data = my_data, 
       paired = TRUE, 
       alternative = "less")
```

```{r}
t.test(weight ~ group, 
       data = my_data, 
       paired = TRUE,
       alternative = "greater")
```

```{r}
# printing the p-value
res$p.value
```

```{r}
# printing the mean
res$estimate
```

```{r}
# printing the confidence interval
res$conf.int
```

## paired samples Wilcoxon test

```{r}
# Weight of the mice before treatment
before <-c(200.1, 190.9, 192.7, 213, 241.4, 196.9, 172.2, 185.5, 205.2, 193.7)
# Weight of the mice after treatment
after <-c(392.9, 393.2, 345.1, 393, 434, 427.9, 422, 383.9, 392.3, 352.2)
# Create a data frame
my_data <- data.frame( 
                group = rep(c("before", "after"), each = 10),
                weight = c(before,  after)
                )
```

```{r}
group_by(my_data, group) %>%
  summarise(
    count = n(),
    median = median(weight, na.rm = TRUE),
    IQR = IQR(weight, na.rm = TRUE)
  )
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800"),
          order = c("before", "after"),
          ylab = "Weight", xlab = "Groups")
```

```{r}
# Subset weight data before treatment
before <- subset(my_data,  group == "before", weight,
                 drop = TRUE)
# subset weight data after treatment
after <- subset(my_data,  group == "after", weight,
                 drop = TRUE)
# Plot paired data
library(PairedData)
pd <- paired(before, after)
plot(pd, type = "profile") + theme_bw()
```

```{r}
res <- wilcox.test(before, after, paired = TRUE)
res
```

```{r}
# Compute t-test
res <- wilcox.test(weight ~ group, data = my_data, paired = TRUE)
res
```

```{r}
# print only the p-value
res$p.value
```

The p-value of the test is 0.001953, which is less than the significance level alpha = 0.05. 
We can conclude that the median weight of the mice before treatment is significantly different from the median weight after treatment with a p-value = 0.001953.

```{r}
wilcox.test(weight ~ group, data = my_data, paired = TRUE, alternative = "less")
```

```{r}
wilcox.test(weight ~ group, data = my_data, paired = TRUE,
       alternative = "greater")
```

## one-way analysis of variance (ANOVA)

```{r}
my_data <- PlantGrowth
```

```{r}
# Show the levels
levels(my_data$group)
```

```{r}
my_data$group <- ordered(my_data$group,
                         levels = c("ctrl", "trt1", "trt2"))
```

```{r}
group_by(my_data, group) %>%
  summarise(
    count = n(),
    mean = mean(weight, na.rm = TRUE),
    sd = sd(weight, na.rm = TRUE)
  )
```

```{r}
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("ctrl", "trt1", "trt2"),
          ylab = "Weight", xlab = "Treatment")
```

```{r}
# Mean plots
# ++++++++++++++++++++
# Plot weight by group
# Add error bars: mean_se
# (other values include: mean_sd, mean_ci, median_iqr, ....)
ggline(my_data, x = "group", y = "weight", 
       add = c("mean_se", "jitter"), 
       order = c("ctrl", "trt1", "trt2"),
       ylab = "Weight", xlab = "Treatment")
```

```{r}
# Box plot
boxplot(weight ~ group, data = my_data,
        xlab = "Treatment", ylab = "Weight",
        frame = FALSE, col = c("#00AFBB", "#E7B800", "#FC4E07"))
```

```{r}
library("gplots")
plotmeans(weight ~ group, data = my_data, frame = FALSE,
          xlab = "Treatment", ylab = "Weight",
          main="Mean Plot with 95% CI") 
```


```{r}
# Compute the analysis of variance
res.aov <- aov(weight ~ group, data = my_data)
# Summary of the analysis
summary(res.aov)
```

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the groups highlighted with “*" in the model summary.

### Tukey multiple pairwise-comparisons

```{r}
TukeyHSD(res.aov)
```

It can be seen from the output, that only the difference between trt2 and trt1 is significant with an adjusted p-value of 0.012.

### multiple pairwise-comparisons

```{r}
library(multcomp)
summary(glht(res.aov, linfct = mcp(group = "Tukey")))
```

### pairewise t-test

```{r}
pairwise.t.test(my_data$weight, my_data$group, p.adjust.method = "BH")
```

### Check ANOVA assumptions

#### Check the homogeneity of variance assumption

```{r}
# 1. Homogeneity of variances
plot(res.aov, 1)
```

In the plot above, there is no evident relationships between residuals and fitted values (the mean of each groups), which is good. 
So, we can assume the homogeneity of variances.

```{r}
library(car)
leveneTest(weight ~ group, data = my_data)
```

From the output above we can see that the p-value is not less than the significance level of 0.05. 
This means that there is no evidence to suggest that the variance across groups is statistically significantly different. 
Therefore, we can assume the homogeneity of variances in the different treatment groups.

#### Relaxing the homogeneity of variance assumption

```{r}
oneway.test(weight ~ group, data = my_data)
```

```{r}
pairwise.t.test(my_data$weight, 
                my_data$group,
                p.adjust.method = "BH", 
                pool.sd = FALSE)
```

#### Check the normality assumption

```{r}
# 2. Normality
plot(res.aov, 2)
```

```{r}
# Extract the residuals
aov_residuals <- residuals(object = res.aov )
# Run Shapiro-Wilk test
shapiro.test(x = aov_residuals)
```

```{r}
kruskal.test(weight ~ group, data = my_data)
```

## Kruskal-Wallis test

```{r}
my_data <- PlantGrowth
```

```{r}
# Show the group levels
levels(my_data$group)
```

```{r}
my_data %>% 
  group_by(group) %>% 
  summarise(
    count = n(),
    mean = mean(weight, na.rm = TRUE),
    sd = sd(weight, na.rm = TRUE),
    median = median(weight, na.rm = TRUE),
    IQR = IQR(weight, na.rm = TRUE)
  )
```

```{r}
# Box plots
# ++++++++++++++++++++
# Plot weight by group and color by group
library("ggpubr")
ggboxplot(my_data, x = "group", y = "weight", 
          color = "group", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("ctrl", "trt1", "trt2"),
          ylab = "Weight", xlab = "Treatment")
```

```{r}
# Mean plots
# ++++++++++++++++++++
# Plot weight by group
# Add error bars: mean_se
# (other values include: mean_sd, mean_ci, median_iqr, ....)
library("ggpubr")
ggline(my_data, x = "group", y = "weight", 
       add = c("mean_se", "jitter"), 
       order = c("ctrl", "trt1", "trt2"),
       ylab = "Weight", xlab = "Treatment")
```

```{r}
kruskal.test(weight ~ group, data = my_data)
```

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the treatment groups.

```{r}
pairwise.wilcox.test(PlantGrowth$weight, 
                     PlantGrowth$group,
                     p.adjust.method = "BH")
```

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
```

## One-way ANOVA

```{r}
data("PlantGrowth")
set.seed(1234)
PlantGrowth

levels(PlantGrowth$group)

PlantGrowth %>%
  group_by(group) %>%
  get_summary_stats(weight, type = "mean_sd")

ggboxplot(PlantGrowth, x = "group", y = "weight")

PlantGrowth %>% 
  group_by(group) %>%
  identify_outliers(weight)

# Build the linear model
model  <- lm(weight ~ group, data = PlantGrowth)

# Create a QQ plot of residuals
ggqqplot(residuals(model))

# Compute Shapiro-Wilk test of normality
shapiro_test(residuals(model))

PlantGrowth %>%
  group_by(group) %>%
  shapiro_test(weight)

ggqqplot(PlantGrowth, "weight", facet.by = "group")

plot(model, 1)

PlantGrowth %>% levene_test(weight ~ group)

# classical one-way ANOVA test
res.aov <- PlantGrowth %>% 
  anova_test(weight ~ group)
res.aov

# Pairwise comparisons
pwc <- PlantGrowth %>% 
  tukey_hsd(weight ~ group)

# Visualization: box plots with p-values
pwc <- pwc %>% 
  add_xy_position(x = "group")
ggboxplot(PlantGrowth, x = "group", y = "weight") +
  stat_pvalue_manual(pwc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc))
```

```{r}
# Welch One way ANOVA test
res.aov2 <- PlantGrowth %>% 
  welch_anova_test(weight ~ group)

# Pairwise comparisons (Games-Howell)
pwc2 <- PlantGrowth %>% 
  games_howell_test(weight ~ group)

# Visualization: box plots with p-values
pwc2 <- pwc2 %>% 
  add_xy_position(x = "group", step.increase = 1)
ggboxplot(PlantGrowth, x = "group", y = "weight") +
  stat_pvalue_manual(pwc2, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.aov2, detailed = TRUE),
    caption = get_pwc_label(pwc2))
```

## Kruskal-Wallis test

```{r}
PlantGrowth

levels(PlantGrowth$group)

PlantGrowth %>% 
  group_by(group) %>%
  get_summary_stats(weight, type = "common")

ggboxplot(PlantGrowth, x = "group", y = "weight")

res.kruskal <- PlantGrowth %>% 
  kruskal_test(weight ~ group)

PlantGrowth %>% 
  kruskal_effsize(weight ~ group)

# Pairwise comparisons
pwc <- PlantGrowth %>% 
  dunn_test(weight ~ group, p.adjust.method = "bonferroni")

pwc2 <- PlantGrowth %>% 
  wilcox_test(weight ~ group, p.adjust.method = "bonferroni")

# Visualization: box plots with p-values
pwc <- pwc %>% 
  add_xy_position(x = "group")
ggboxplot(PlantGrowth, x = "group", y = "weight") +
  stat_pvalue_manual(pwc, hide.ns = TRUE, y.position = c(7, 8, 9)) +
  labs(subtitle = get_test_label(res.kruskal, detailed = TRUE),
       caption = get_pwc_label(pwc))
```


```{r}
library(rstatix)
library(ggpubr)

data("PlantGrowth")

# conduct Kruskal Wallis test
res.kruskal <- PlantGrowth %>% 
  kruskal_test(weight ~ group)

# conduct pairwise comparison
pwc <- PlantGrowth %>% 
  dunn_test(weight ~ group, p.adjust.method = "bonferroni")

# Visualization: box plots with p-values
pwc <- pwc %>% 
  add_xy_position(x = "group", fun = "median")

ggboxplot(PlantGrowth, x = "group", y = "weight") +
  stat_pvalue_manual(pwc) +
  labs(subtitle = get_test_label(res.kruskal, detailed = TRUE),
       caption = get_pwc_label(pwc))
```
```{r}
reprex()
```




## ANCOVA

```{r}
library(broom)
library(datarium)
```

```{r}
# Load and prepare the data
data("anxiety", package = "datarium")
anxiety <- anxiety %>%
  select(id, group, t1, t3) %>%
  rename(pretest = t1, posttest = t3)
anxiety[14, "posttest"] <- 19

anxiety
```

```{r}
ggscatter(anxiety, x = "pretest", y = "posttest", color = "group", add = "reg.line")+
  stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), 
                            color = group))

anxiety %>% 
  anova_test(posttest ~ group*pretest)

# Fit the model, the covariate goes first
model <- lm(posttest ~ pretest + group, data = anxiety)
# Inspect the model diagnostic metrics
model.metrics <- augment(model) %>%
  select(-.hat, -.sigma, -.fitted, -.se.fit) # Remove details
head(model.metrics, 3)

# Assess normality of residuals using shapiro wilk test
shapiro_test(model.metrics$.resid)

model.metrics %>% levene_test(.resid ~ group)

model.metrics %>% 
  filter(abs(.std.resid) > 3) %>%
  as.data.frame()

res.aov <- anxiety %>% 
  anova_test(posttest ~ pretest + group)
get_anova_table(res.aov)

# Pairwise comparisons
library(emmeans)
pwc <- anxiety %>% 
  emmeans_test(
    posttest ~ group, covariate = pretest,
    p.adjust.method = "bonferroni"
    )
pwc

# Display the adjusted means of each group
# Also called as the estimated marginal means (emmeans)
get_emmeans(pwc)

```

```{r}
# Visualization: line plots with p-values
pwc <- pwc %>% add_xy_position(x = "group", fun = "mean_se")
ggline(get_emmeans(pwc), x = "group", y = "emmean") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) + 
  stat_pvalue_manual(pwc, hide.ns = TRUE, tip.length = FALSE) +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
  )
```

```{r}
data("stress", package = "datarium")
stress

# check linearity assumption
ggscatter(stress, x = "age", y = "score",
          facet.by  = c("exercise", "treatment"), 
          short.panel.labs = FALSE) +
  stat_smooth(method = "loess", span = 0.9)

# check homogeneity of regression slopes
stress %>%
  anova_test(score ~ age + treatment + exercise + 
               treatment * exercise + age * treatment + age * exercise + 
               age * exercise * treatment)

stress %>%
  unite(col = "group", treatment, exercise) %>%
  anova_test(score ~ group*age)

# check normality of residuals
## Fit the model, the covariate goes first
model <- lm(score ~ age + treatment * exercise, data = stress)

## Inspect the model diagnostic metrics
model.metrics <- augment(model) %>%
  select(-.hat, -.sigma, -.fitted, -.se.fit) # Remove details

## Assess normality of residuals using shapiro wilk test
shapiro_test(model.metrics$.resid)

# check homogeneity of variances
levene_test(.resid ~ treatment * exercise, data = model.metrics)

# check outliers
model.metrics %>% 
  filter(abs(.std.resid) > 3) %>%
  as.data.frame()

# Computation
res.aov <- stress %>% 
  anova_test(score ~ age + treatment * exercise)
get_anova_table(res.aov)

# Effect of treatment at each level of exercise
stress %>%
  group_by(exercise) %>%
  anova_test(score ~ age + treatment)

# Pairwise comparisons
pwc <- stress %>% 
  group_by(exercise) %>%
  emmeans_test(score ~ treatment, 
               covariate = age,
               p.adjust.method = "bonferroni")
pwc %>% filter(exercise == "high")

# Effect of exercise at each level of treatment
stress %>%
  group_by(treatment) %>%
  anova_test(score ~ age + exercise)

pwc2 <- stress %>% 
  group_by(treatment) %>%
  emmeans_test(score ~ exercise, 
               covariate = age,
               p.adjust.method = "bonferroni") %>%
  select(-df, -statistic, -p) # Remove details
pwc2 %>% filter(treatment == "yes")

# Line plot
lp <- ggline(get_emmeans(pwc), x = "exercise", y = "emmean", 
             color = "treatment", palette = "jco") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = treatment), width = 0.1)

# Comparisons between treatment group at each exercise level
pwc <- pwc %>% 
  add_xy_position(x = "exercise", fun = "mean_se", step.increase = 0.2)

pwc.filtered <- pwc %>% 
  filter(exercise == "high")

lp + 
  stat_pvalue_manual(pwc.filtered, hide.ns = TRUE, tip.length = 0, bracket.size = 0) +
  labs(subtitle = get_test_label(res.aov,  detailed = TRUE), caption = get_pwc_label(pwc))

# Comparisons between exercises group at each treatment level
pwc2 <- pwc2 %>% add_xy_position(x = "exercise", fun = "mean_se")
pwc2.filtered <- pwc2 %>% filter(treatment == "yes")
lp + 
  stat_pvalue_manual(pwc2.filtered, hide.ns = TRUE, tip.length = 0,
                   step.group.by = "treatment", color = "treatment") +
  labs(subtitle = get_test_label(res.aov,  detailed = TRUE), caption = get_pwc_label(pwc2))
```

## rstatix

```{r}
library(rstatix)  
library(ggpubr)  # For easy data-visualization
```

```{r}
df <- ToothGrowth
df$dose <- as.factor(df$dose)

df %>% 
  t_test(len ~ 1, mu = 0)

# One-sample test of each dose level
df %>% 
  group_by(dose) %>%
  t_test(len ~ 1, mu = 0)

# T-test
stat.test <- df %>% 
  t_test(len ~ supp, paired = FALSE) 
stat.test
```


```{r}
# Create a box plot
ggboxplot(df, x = "supp", y = "len", color = "supp", palette = "jco", ylim = c(0, 40)) +
  stat_pvalue_manual(stat.test, label = "T-test, p = {p}", y.position = 35)
```


```{r}
# Statistical test
stat.test <- df %>%
  group_by(dose) %>%
  t_test(len ~ supp) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test
```

```{r}
ggboxplot(df, x = "supp", y = "len",
          color = "supp", palette = "jco", facet.by = "dose",ylim = c(0, 40)) +
  stat_pvalue_manual(stat.test, label = "p.adj", y.position = 35)
```

```{r}
# T-test
stat.test <- df %>% 
  t_test(len ~ supp, paired = TRUE) 
stat.test
```

```{r}
# Box plot
ggpaired(df, x = "supp", y = "len", color = "supp", palette = "jco", 
         line.color = "gray", line.size = 0.4, ylim = c(0, 40)) +
  stat_pvalue_manual(stat.test, label = "p", y.position = 36)
```

```{r}
# Pairwise t-test
pairwise.test <- df %>% 
  t_test(len ~ dose)
pairwise.test
```

```{r}
# Box plot
ggboxplot(df, x = "dose", y = "len")+
  stat_pvalue_manual(pairwise.test, label = "p.adj", y.position = c(29, 35, 39))
```

```{r}
# Comparison against reference group
#::::::::::::::::::::::::::::::::::::::::
# T-test: each level is compared to the ref group
stat.test <- df %>% t_test(len ~ dose, ref.group = "0.5")
stat.test
```

```{r}
# Box plot
ggboxplot(df, x = "dose", y = "len", ylim = c(0, 40)) +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", y.position = c(29, 35))
```

```{r}
# T-test
stat.test <- df %>% t_test(len ~ dose, ref.group = "all")
stat.test
```

```{r}
# Box plot with horizontal mean line
ggboxplot(df, x = "dose", y = "len") +
  stat_pvalue_manual(stat.test, label = "p.adj.signif", y.position = 35, remove.bracket = TRUE) +
  geom_hline(yintercept = mean(df$len), linetype = 2)
```

## Add P-values

```{r}
library(ggpubr)

data("ToothGrowth")
ToothGrowth
```


```{r}
compare_means(len ~ supp, data = ToothGrowth)
```

```{r}
ggboxplot(ToothGrowth, x = "supp", y = "len", color = "supp", 
          palette = "jco", add = "jitter") +
  stat_compare_means(label.x = 1, label.y = 33) +
  stat_compare_means(method = "t.test", label.x = 1, label.y = 36) +
  stat_compare_means(aes(label = ..p.signif..),  label.x = 1.5, label.y = 40) + 
  stat_compare_means( label = "p.signif", label.x = 2, label.y = 40)
```

```{r}
compare_means(len ~ supp, data = ToothGrowth, paired = TRUE)
```

```{r}
ggpaired(ToothGrowth, x = "supp", y = "len",
         color = "supp", line.color = "gray", line.size = 0.4, palette = "jco") +
  stat_compare_means(paired = TRUE)
```

```{r}
# Global test
compare_means(len ~ dose,  data = ToothGrowth, method = "anova")
```


```{r}
# Default method = "kruskal.test" for multiple groups
ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "dose", palette = "jco") +
  stat_compare_means()

# Change method to anova
ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "dose", palette = "jco") +
  stat_compare_means(method = "anova")
```


```{r}
# Perorm pairwise comparisons
compare_means(len ~ dose,  data = ToothGrowth)
```

```{r}
# Visualize: Specify the comparisons you want
my_comparisons <- list( c("0.5", "1"), c("1", "2"), c("0.5", "2") )

ggboxplot(ToothGrowth, x = "dose", y = "len", color = "dose", palette = "jco")+ 
  stat_compare_means(comparisons = my_comparisons) + # Add pairwise comparisons p-value
  stat_compare_means(label.y = 50)     # Add global p-value

ggboxplot(ToothGrowth, x = "dose", y = "len", color = "dose", palette = "jco")+ 
  stat_compare_means(comparisons = my_comparisons, label.y = c(29, 35, 40)) +
  stat_compare_means(label.y = 45)
```

```{r}
# Pairwise comparison against reference
compare_means(len ~ dose, data = ToothGrowth, ref.group = "0.5", method = "t.test")
```

```{r}
# Visualize
ggboxplot(ToothGrowth, x = "dose", y = "len", color = "dose", palette = "jco") +
  stat_compare_means(method = "anova", label.y = 40)+      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test", ref.group = "0.5")                    # Pairwise comparison against reference
```

```{r}
# Comparison of each group against base-mean
compare_means(len ~ dose,  data = ToothGrowth, ref.group = ".all.",
              method = "t.test")
```

```{r}
# Visualize
ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "dose", palette = "jco")+
  stat_compare_means(method = "anova", label.y = 40)+      # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")
```

```{r}
# Load myeloma data from GitHub
myeloma <- read.delim("https://raw.githubusercontent.com/kassambara/data/master/myeloma.txt")
# Perform the test
compare_means(DEPDC1 ~ molecular_group,  data = myeloma, ref.group = ".all.", method = "t.test")
```

```{r}
# Visualize the expression profile
ggboxplot(myeloma, x = "molecular_group", y = "DEPDC1", 
          color = "molecular_group", add = "jitter", legend = "none") +
  rotate_x_text(angle = 45)+
  # Add horizontal line at base mean
  geom_hline(yintercept = mean(myeloma$DEPDC1), linetype = 2) +
  # Add global annova p-value
  stat_compare_means(method = "anova", label.y = 1600) +
  # Pairwise comparison against all
  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.")

# Visualize the expression profile
ggboxplot(myeloma, x = "molecular_group", y = "DEPDC1", 
          color = "molecular_group", add = "jitter", legend = "none") +
  rotate_x_text(angle = 45) +
  # Add horizontal line at base mean
  geom_hline(yintercept = mean(myeloma$DEPDC1), linetype = 2) + 
  # Add global annova p-value
  stat_compare_means(method = "anova", label.y = 1600) + 
  # Pairwise comparison against all
  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE)      
```




```{r}
compare_means(len ~ supp, data = ToothGrowth, group.by = "dose")
```

```{r}
# Box plot facetted by "dose"
p <- ggboxplot(ToothGrowth, x = "supp", y = "len", 
               color = "supp", palette = "jco", add = "jitter",
               facet.by = "dose", short.panel.labs = FALSE)

# Use only p.format as label. Remove method name.
p + stat_compare_means(label = "p.format")

# Or use significance symbol as label
p + stat_compare_means(label =  "p.signif", label.x = 1.5)
```

```{r}
p <- ggboxplot(ToothGrowth, x = "dose", y = "len", color = "supp", 
               palette = "jco", add = "jitter")
p + stat_compare_means(aes(group = supp))

# Show only p-value
p + stat_compare_means(aes(group = supp), label = "p.format")

# Use significance symbol as label
p + stat_compare_means(aes(group = supp), label = "p.signif")
```

```{r}
compare_means(len ~ supp, data = ToothGrowth, group.by = "dose", paired = TRUE)
```

```{r}
# Box plot facetted by "dose"
p <- ggpaired(ToothGrowth, x = "supp", y = "len", color = "supp", 
              palette = "jco", line.color = "gray", line.size = 0.4,
              facet.by = "dose", short.panel.labs = FALSE)

# Use only p.format as label. Remove method name.
p + stat_compare_means(label = "p.format", paired = TRUE)
```

```{r}
# Bar plot of mean +/-se
ggbarplot(ToothGrowth, x = "dose", y = "len", add = "mean_se") +
  # Global p-value
  stat_compare_means() +
  # compare to ref.group
  stat_compare_means(ref.group = "0.5", label = "p.signif", label.y = c(22, 29))

# Line plot of mean +/-se
ggline(ToothGrowth, x = "dose", y = "len", add = "mean_se") +
  # Global p-value
  stat_compare_means() +                                         
  stat_compare_means(ref.group = "0.5", label = "p.signif", label.y = c(22, 29)) 
```

```{r}
ggbarplot(ToothGrowth, x = "dose", y = "len", add = "mean_se",
          color = "supp", palette = "jco", 
          position = position_dodge(0.8))+
  stat_compare_means(aes(group = supp), label = "p.signif", label.y = 29)

ggline(ToothGrowth, x = "dose", y = "len", add = "mean_se",
          color = "supp", palette = "jco")+
  stat_compare_means(aes(group = supp), label = "p.signif", 
                     label.y = c(16, 25, 29))
```

## robustt ANCOVA

```{r}
library(WRS2)
summary(invisibility)

## specifying covariate evaluation points
ancova(mischief2 ~ cloak + mischief1, data = invisibility, pts = c(3, 4, 8, 1))

## bootstrap version
ancboot(mischief2 ~ cloak + mischief1, data = invisibility)
```

## Normality Test in R

```{r}
summary(ToothGrowth)

# Density plot
ggdensity(ToothGrowth$len, fill = "lightgray")

# QQ plot
ggqqplot(ToothGrowth$len)

ToothGrowth %>% 
  shapiro_test(len)

ToothGrowth %>%
  group_by(dose) %>%
  shapiro_test(len)

iris %>% 
  shapiro_test(Sepal.Length, Petal.Width)
```


## Homogeneity of Variance Test in R

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

var.test(len ~ supp, data = ToothGrowth)

bartlett.test(weight ~ group, data = PlantGrowth)

bartlett.test(len ~ interaction(supp,dose), data=ToothGrowth)

library(car)
# Levene's test with one independent variable
leveneTest(weight ~ group, data = PlantGrowth)

# Levene's test with multiple independent variables
leveneTest(len ~ supp * dose, data = ToothGrowth)

fligner.test(weight ~ group, data = PlantGrowth)
```

## Transform Data to Normal Distribution

```{r}
library(moments)
skewness(iris$Sepal.Length, na.rm = TRUE)
```

```{r}
data("USJudgeRatings")
df <- USJudgeRatings

# Distribution of CONT variable
ggdensity(df, x = "CONT", fill = "lightgray", title = "CONT") +
  scale_x_continuous(limits = c(3, 12)) +
  # Overlay normal density plot (with the same mean and SD) to the density distribution of 'x'.
  stat_overlay_normal_density(color = "red", linetype = "dashed")

skewness(df$CONT, na.rm = TRUE)

df$CONT <- log10(df$CONT)

# Distribution of CONT variable
ggdensity(df, x = "CONT", fill = "lightgray", title = "CONT") +
  stat_overlay_normal_density(color = "red", linetype = "dashed")

skewness(df$CONT, na.rm = TRUE)

# Distribution of PHYS variable
ggdensity(df, x = "PHYS", fill = "lightgray", title = "PHYS") +
  scale_x_continuous(limits = c(3, 12)) +
  stat_overlay_normal_density(color = "red", linetype = "dashed")

skewness(df$PHYS, na.rm = TRUE)

df$PHYS <- log10(max(df$CONT+1) - df$CONT)

# Distribution of PHYS variable
ggdensity(df, x = "PHYS", fill = "lightgray", title = "PHYS") +
  stat_overlay_normal_density(color = "red", linetype = "dashed")

skewness(df$PHYS, na.rm = TRUE)
```

## One-Way MANOVA

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(car)
library(broom)
```

```{r}
iris2 <- iris %>%
  select(Sepal.Length, Petal.Length, Species) %>%
  add_column(id = 1:nrow(iris), .before = 1)

ggboxplot(iris2, x = "Species", y = c("Sepal.Length", "Petal.Length"), 
          merge = TRUE, palette = "jco")

iris2 %>%
  group_by(Species) %>%
  get_summary_stats(Sepal.Length, Petal.Length, type = "mean_sd")

iris2 %>%
  group_by(Species) %>%
  summarise(N = n())

iris2 %>%
  group_by(Species) %>%
  identify_outliers(Sepal.Length)

iris2 %>%
  group_by(Species) %>%
  identify_outliers(Petal.Length)

iris2 %>%
 group_by(Species) %>%
 mahalanobis_distance(-id) %>%
 filter(is.outlier == TRUE)

iris2 %>%
  group_by(Species) %>%
  shapiro_test(Sepal.Length, Petal.Length) %>%
  arrange(variable)

# QQ plot of Sepal.Length
ggqqplot(iris2, "Sepal.Length", facet.by = "Species",
         ylab = "Sepal Length", ggtheme = theme_bw())

# QQ plot of Petal.Length
ggqqplot(iris2, "Petal.Length", facet.by = "Species",
         ylab = "Petal Length", ggtheme = theme_bw())
```

```{r}
iris2 %>%
  select(Sepal.Length, Petal.Length) %>%
  mshapiro_test()

iris2 %>% cor_test(Sepal.Length, Petal.Length)

library(GGally)
results <- iris2 %>%
  select(Sepal.Length, Petal.Length, Species) %>%
  group_by(Species) %>%
  doo(~ggpairs(.) + theme_bw(), result = "plots")
results$plots

box_m(iris2[, c("Sepal.Length", "Petal.Length")], iris2$Species)

iris2 %>% 
  gather(key = "variable", value = "value", Sepal.Length, Petal.Length) %>%
  group_by(variable) %>%
  levene_test(value ~ Species)

model <- lm(cbind(Sepal.Length, Petal.Length) ~ Species, iris2)
Manova(model, test.statistic = "Pillai")

```

