---
title: "Leaflet"
author: "Xiaochi"
date: "23/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)

library(leaflet)
library(rgeos)
```

# Basic Usage

1. Create a map widget by calling leaflet( ).
2. Add layers (features) to the map by using layer functions
    + addTiles
    + addMarkers
    + addPolygons
3. Repeat step 2 as desired.
4. Print the map widget to display it.

```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = 174.768, lat = -36.852, popup = "The birthplace of R")
```

# The Map Widget

The function leaflet( ) returns a Leaflet map widget, which stores a list of objects that can be modified or updated later.

Most functions in this package have an argument map as their first argument, which makes it easy to use the pipe operator %>% .

## Initializing Options

```{r}
leaflet(options = leafletOptions(minZoom = 0, maxZoom = 3))
```

Using leafletOptions( ) can set a custom CRS.

## Map Methods

* setView( ) sets the center of the map view and the zoom level;
* fitBounds() fits the view into the rectangle [lng1, lat1] – [lng2, lat2];
* clearBounds() clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided;

## The Data Object

Both leaflet() and the map layer functions have an optional data parameter that is designed to receive spatial data in one of several forms:

* From base R:
    * lng/lat matrix
    * data frame with lng/lat columns

* From the sp package:
    * SpatialPoints[DataFrame]
    * Line/lines
    * SpatialLines[DataFrame]
    * Polygon/Polygons
    * SpatialPolygons[DataFrame]

* From the maps package:
    * the data frame from returned from map( )

The data argument is used to derive spatial data for functions that need it; for example, if data is a SpatialPolygonsDataFrame object, then calling addPolygons on that map widget will know to add the polygons from that SpatialPolygonsDataFrame.

It is straightforward to derive these variables from sp objects since they always represent spatial data in the same way. On the other hand, for a normal matrix or data frame, any numeric column could potentially contain spatial data. So we resort to guessing based on column names:

* latitude: lat, latitude
* longtitude: lng, long, longtitude
    
```{r}
df = data.frame(Lat = 1:10, Long = rnorm(10))
df
leaflet(df) %>% 
  addCircles()
```

```{r}
leaflet(df) %>% 
  addCircles(lng = ~Long, lat = ~Lat)
```

```{r}
leaflet() %>% 
  addCircles(data = df)
```

```{r}
leaflet() %>% 
  addCircles(data = df, lat = ~ Lat, lng = ~ Long)
```

```{r}
library(sp)
Sr1 = Polygon(cbind(c(2, 4, 4, 1, 2),
                    c(2, 3, 5, 4, 2)))
Sr2 = Polygon(cbind(c(5, 4, 2, 5),
                    c(2, 3, 2, 2)))
Sr3 = Polygon(cbind(c(4, 4, 5, 10, 4), 
                    c(5, 3, 2, 5, 5)))
Sr4 = Polygon(cbind(c(5, 6, 6, 5, 5), 
                    c(4, 4, 3, 3, 4)), 
              hole = TRUE)
Srs1 = Polygons(list(Sr1), "s1")
Srs2 = Polygons(list(Sr2), "s2")
Srs3 = Polygons(list(Sr4, Sr3), "s3/4")
SpP = SpatialPolygons(list(Srs1, Srs2, Srs3), 1:3)
leaflet(height = "300px") %>% addPolygons(data = SpP)
```

```{r}
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% 
  addTiles() %>% 
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```

## The Formula Interface

The arguments of all layer functions can take normal R objects, such as a numeric vector for the lat argument, or a character vector of colors for the color argument. 

They can also take a one-sided formula, in which case the formula will be evaluated using the "data" argument as the environment. 

For example, ~ x means the variable x in the data object, and you can write arbitrary expressions on the right-hand side, e.g., ~ sqrt(x + 1).

```{r}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)

leaflet(df) %>% 
  addTiles() %>% 
  addCircleMarkers(radius = ~size, color = ~color, fill = FALSE) %>% 
  addCircleMarkers(radius = runif(100, 4, 10), color = c("red"))
```

# Using Basemaps

## Default (OpenStreetMap) Tiles

addTiles( )

```{r}
m <- leaflet() %>% 
  setView(lng = -71.0589, lat = 42.3601, zoom = 12) %>% 
  addTiles()
```

## Thifrd-Party Tiles

addProviderTiles( )

```{r}
m %>% addProviderTiles(providers$Stamen.Toner)
```

```{r}
m %>% addProviderTiles(providers$CartoDB.Positron)
```

```{r}
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```


## Combining Tile Layers

```{r}
m %>% 
  addProviderTiles(providers$MtbMap) %>% 
  addProviderTiles(providers$Stamen.TonerLines, 
                   options = providerTileOptions(opacity = 0.35)) %>% 
  addProviderTiles(providers$Stamen.TonerLabels)

?setView
```

# Markers

addMarkers( )
addAwesomeMarkers( )

* popup: add a message to be displayed on click.
* label: display a text label either on hover or statically 

```{r}
data(quakes)

leaflet(data = quakes[1:20,]) %>% 
  addTiles() %>% 
  addAwesomeMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

## Customizing Marker Icons

### makeIcon( )

applying a single icon to a set of markers

```{r}
greenLeafIcon <- makeIcon(
  iconUrl = "http://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) %>% 
  addTiles() %>% 
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

### icons( )

several icons vary only by a couple of parameters (i.e. they share the same size and anchor points but have different URLs) 

```{r}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "http://leafletjs.com/examples/custom-icons/leaf-green.png",
    "http://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) %>% 
  addTiles() %>% 
  addMarkers(~long, ~lat, icon = leafIcons)


```

### iconList( )

have a set of icons that vary in multiple parameters.

iconList() can create a list of makeIcon() icons, and select rom that list by position or name.

```{r}
oceanIcons <- iconList(
  ship = makeIcon("ferry-18.png", "ferry-18@2x.png", 18, 18),
  pirate = makeIcon("danger-24.png", "danger-24@2x.png", 24, 24)
)

df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 90.620130,  # lng
    (runif(20) - .5) * 3.8 + 25.638077  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "pirate", "ship"),
    c("ship", "pirate")
  ))
)

leaflet(df) %>% 
  addTiles() %>% 
  addMarkers(icon = ~oceanIcons[type])
```

## Awesome Icons

* makeAwesomeIcon
* awesomeIcons
* awesomeIconList

```{r}
df.20 <- quakes[1:20, ]

getColor <- function(quakes) {
  sapply(quakes$mag, function(mag) {
    if (mag <= 4) {
      "green"
    } else if (mag <= 5) {
      "orange"
    } else {
      "red"
    }
  })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(df.20)
)

leaflet(df.20) %>%
  addTiles() %>%
  addAwesomeMarkers( ~ long,
                     ~ lat,
                     icon = icons,
                     label = ~ as.character(mag))
```

## Marker Clusters

```{r}
leaflet(quakes) %>% 
  addTiles() %>% 
  addMarkers(clusterOptions = markerClusterOptions())
```


## Circle Markers

```{r}
leaflet(df) %>% 
  addTiles() %>% 
  addCircleMarkers()
```

```{r}
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )
```

# Popups

addPopups( ): add standalone popup to the map

```{r}
content <- paste(sep = "<br/>",
  "<b><a href='http://www.samurainoodle.com'>Samurai Noodle</a></b>",
  "606 5th Ave. S",
  "Seattle, WA 98138"
)

leaflet() %>% 
  addTiles() %>% 
  addPopups(-122.327298, 47.597131, content,
            options = popupOptions(closeButton = FALSE))
```

```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"
))

leaflet(df) %>% 
  addTiles() %>% 
  addMarkers(~Long, ~Lat, popup = ~htmlEscape(Name))
```

# Labels

always displayed or displayed on mouse over.

Unlike popups you don't need to click a marker/polygon for the label to be shown.

```{r}
library(htmltools)

df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"))

leaflet(df) %>% 
  addTiles() %>% 
  addMarkers(~Long, ~Lat, label = ~htmlEscape(Name))
```

## Customizing Marker Labels

labelOptions argument

```{r}
leaflet() %>% 
  addTiles() %>% 
  setView(-118.456554, 34.09, 13) %>% 
  addMarkers(
    lng = -118.456554, lat = 34.105,
    label = "Default Label",
    labelOptions = labelOptions(noHide = T)
  ) %>% 
  addMarkers(
    lng = -118.456554, lat = 34.095,
    label = "Label w/o surrounding box",
    labelOptions = labelOptions(noHide = T, textOnly = TRUE)
  ) %>% 
  addMarkers(
    lng = -118.456554, lat = 34.085,
    label = "label w/ textsize 15px",
    labelOptions = labelOptions(noHide = T, textsize = "15px")
  ) %>% 
  addMarkers(
    lng = -118.456554, lat = 34.075,
    label = "Label w/ custom CSS style",
    labelOptions = labelOptions(noHide = T, direction = "bottom",
      style = list(
        "color" = "red",
        "font-family" = "serif",
        "font-style" = "italic",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"
      )))
  
```

# Colors

```{r}
pal <- colorNumeric(c("red", "green", "blue"), 1:10)
pal(c(1,6,9))
```

