---
title: "Regression"
author: "Xiaochi"
date: "23/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# OLS regression

## Simple linear regression

```{r}
women
```

```{r}
fit <- lm(weight ~ height, data = women)
summary(fit)
```

the regression coefficient 3.45 is significantly different from zero (p < 0.001) and indicates that there's an expected increase of 3.45 pounds of weight for every 1 inch increase in height.


```{r}
women$weight
```

```{r}
fitted(fit)
```

```{r}
residuals(fit)
```

```{r}
plot(women$height, 
     women$weight,
     xlab = "Height (in inches)",
     ylab = "Weight (in pounds)")
abline(fit)
```

## Polynomial regression

```{r}
fit2 <- lm(weight ~ height + I(height^2), data = women)
summary(fit2)
```

```{r}
plot(women$height,women$weight,
     xlab="Height (in inches)",
     ylab="Weight (in lbs)")
lines(women$height, fitted(fit2))
```

```{r}
library(car)
scatterplot(weight ~ height, 
            data=women,
            spread=FALSE, 
            smoother.args=list(lty=2), 
            pch=19,
            main="Women Age 30-39",
            xlab="Height (inches)",
            ylab="Weight (lbs.)")
```

## Multiple linear regression

```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population", "Illiteracy", "Income", "Frost")])
```

```{r}
cor(states)
```

```{r}
library(car)
scatterplotMatrix(states,main="Scatter Plot Matrix")
```

```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population", "Illiteracy", "Income", "Frost")])

fit <- lm(Murder ~ Population + Illiteracy + Income + Frost,
          data=states)

summary(fit)
```

the regression coefficient for Illiteracy is 4.14, suggesting that an increase of 1% in illiteracy is associated with a 4.14% increase in the murder rate, controlling for population, income, and temperature.

```{r}
mtcars
```

```{r}
fit <- lm(mpg ~ hp + wt + hp:wt, data = mtcars)
summary(fit)
```

```{r}
library(effects)
plot(effect("hp:wt", fit, , list(wt = c(2.2, 3.2, 4.2))), multiline=TRUE)
```

# Regression diagnostics

```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population", "Illiteracy", "Income", "Frost")])

fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
```


```{r}
confint(fit)
```

The results suggest that you can be 95% confident that the interval [2.38, 5.90] contains the true change in murder rate for a 1% change in illiteracy rate. Additionally, because the confidence interval for Frost contains 0, you can conclude that a change in temperature is unrelated to murder rate, holding the other variables constant.

## A typical approach

```{r}
fit <- lm(weight ~ height, data=women)
par(mfrow=c(2,2))
plot(fit)
```

```{r}
fit2 <- lm(weight ~ height + I(height^2), data=women)
par(mfrow=c(2,2))
plot(fit2)
```

```{r}
newfit <- lm(weight~ height + I(height^2), data=women[-c(13,15),])
```

```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population", "Illiteracy", "Income", "Frost")])

fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)

par(mfrow=c(2,2))
plot(fit)
```

## An enhanced approach

```{r}
library(car)

states <- as.data.frame(state.x77[,c("Murder", "Population", "Illiteracy", "Income", "Frost")])

fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)

qqPlot(fit, 
       labels=row.names(states), 
       id.method="identify",
       simulate=TRUE, 
       main="Q-Q Plot")
```

```{r}
states["Nevada",]
```

```{r}
fitted(fit)["Nevada"]
```

```{r}
residuals(fit)["Nevada"]
```

```{r}
rstudent(fit)["Nevada"]
```

```{r}
residplot <- function(fit, nbreaks=10) {
        z <- rstudent(fit)
        hist(z, breaks=nbreaks, 
             freq=FALSE,
             xlab="Studentized Residual",
             main="Distribution of Errors")
        rug(jitter(z), 
            col="brown")
        curve(dnorm(x, mean=mean(z), sd=sd(z)),
              add=TRUE, 
              col="blue", 
              lwd=2)
        lines(density(z)$x, 
              density(z)$y,
              col="red", 
              lwd=2, 
              lty=2)
        legend("topright",
               legend = c( "Normal Curve", "Kernel Density Curve"),
               lty=1:2, 
               col=c("blue","red"), 
               cex=.7)
}

residplot(fit)
```

