---
title: "Feature engineering with recipes"
author: "Xiaochi"
date: "15/03/2021"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidymodels)

data(ames)
ames %>% 
  mutate(Sale_Price = log10(Sale_Price)) -> ames

set.seed(123)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)


lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))

lm_fit <- fit(lm_wflow, ames_train)
```




## A SIMPLE RECIPE FOR THE AMES HOUSING DATA

```{r}
lm(Sale_Price ~ Neighborhood + log10(Gr_Liv_Area) + Year_Built + Bldg_Type, data = ames)
```

```{r}
ames %>% 
  select(Neighborhood, Bldg_Type)
```

```{r}
recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
       data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>%
  step_dummy(all_nominal_predictors()) -> simple_ames

simple_ames
```









## USING RECIPES

```{r}
lm_wflow
```


```{r}
lm_wflow %>% 
  add_recipe(simple_ames)
```


```{r}
lm_wflow %>% 
  remove_variables() %>% 
  add_recipe(simple_ames) -> lm_wflow
lm_wflow
```

```{r}
lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
predict(lm_fit, ames_test %>% slice(1:3))
```


```{r}
# Get the recipe after it has been estimated:
lm_fit %>% 
  extract_recipe(estimated = TRUE)
```

```{r}
# To tidy the model fit: 
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy() %>% 
  slice(1:5)
```


## HOW DATA ARE USED BY THE recipe()


## EXAMPLES OF RECIPE STEPS





### ENCODING QUALITATIVE DATA IN A NUMERIC FORMAT


```{r}
ggplot(ames_train, aes(y = Neighborhood)) + 
  geom_bar() + 
  labs(y = NULL)
```


```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal())

simple_ames
```





### INTERACTION TERMS

```{r}
ames_train %>% 
  select(Gr_Liv_Area)
```


```{r}
ggplot(ames_train, aes(x = Gr_Liv_Area, y = 10^Sale_Price)) + 
  geom_point(alpha = .2) + 
  facet_wrap(~ Bldg_Type) + 
  geom_smooth(method = lm, formula = y ~ x, se = FALSE, col = "red") + 
  scale_x_log10() + 
  scale_y_log10() + 
  labs(x = "General Living Area", y = "Sale Price (USD)")
```

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  # Gr_Liv_Area is on the log scale from a previous step
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") )
```




### SPLINE FUNCTIONS

```{r}
library(patchwork)
library(splines)

plot_smoother <- function(deg_free) {
  ggplot(ames_train, aes(x = Latitude, y = Sale_Price)) + 
    geom_point(alpha = .2) + 
    scale_y_log10() +
    geom_smooth(method = lm,
                formula = y ~ ns(x, df = deg_free),
                col = "red",
                se = FALSE) +
    ggtitle(paste(deg_free, "Spline Terms"))
}

( plot_smoother(2) + plot_smoother(5) ) / ( plot_smoother(20) + plot_smoother(100) )
```

```{r}
recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact(~ Gr_Liv_Area : starts_with("Bldg_Type_")) %>% 
  step_ns(Latitude, deg_free = 20)
```




### FEATURE EXTRACTION
### ROW SAMPLING STEPS
### GENERAL TRANSFORMATIONS
### NATURAL LANGUAGE PROCESSING





## SKIPPING STEPS FOR NEW DATA


## TIDY A recipe()


```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + 
           Latitude + Longitude, data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)

ames_rec
```


```{r}
tidy(ames_rec)
```


```{r}
ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + Latitude + Longitude,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01, id = "my_id") %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)
```


```{r}
lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)
```

```{r}
estimated_recipe <- 
  lm_fit %>% 
  extract_recipe(estimated = TRUE)

tidy(estimated_recipe)
tidy(estimated_recipe, id = "my_id")
```


```{r}
tidy(estimated_recipe, number = 2)
```



## COLUMN ROLES




## CHAPTER SUMMARY

```{r}
library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

ames_rec <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type + 
           Latitude + Longitude, data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_other(Neighborhood, threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact( ~ Gr_Liv_Area:starts_with("Bldg_Type_") ) %>% 
  step_ns(Latitude, Longitude, deg_free = 20)

lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)
```


