---
title: "Graphics for Communication"
author: "Xiaochi"
date: "06/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
library(tidyverse)
```

# Label

labs( ): add labels

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov",
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
    )
```

* title: the purpose of a plot title is to summarise the main finding. Avoid titles that just describe what the plot is, e.g. “A scatterplot of engine displacement vs. fuel economy”.

* subtitle: adds additional detail in a smaller font beneath the title.

* caption: adds text at the bottom right of the plot, often used to describe the source of the data.

* x

* y

* colour


```{r}
df <- tibble(
  x = runif(10),
  y = runif (10)
)

ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + bete + frac(delta, theta))
  )
```

# Annotations

* geom_text( )
    * label
    * data: have a tibble that provides labels

```{r}
best_in_class <- mpg %>% 
  group_by(class) %>% 
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_text(aes(label = model), data = best_in_class)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
```

```{r}
class_avg <- mpg %>% 
  group_by(class) %>% 
  summarise(
    displ = median(displ),
    hwy = median(hwy)
  )

ggplot(mpg, aes(displ, hwy, colour = class)) +
  ggrepel::geom_label_repel(aes(label = class),
                            data = class_avg,
                            size = 6,
                            label.size = 0,
                            segment.color = NA) +
  geom_point() +
  theme(legend.position = "none")

```

```{r}
label <- mpg %>% 
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \n related to decreasing fuel economy."
  )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "rigth")
```

```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")

```

```{r}
"Increasing engine size is related to decreasing fuel economy." %>% 
  stringr::str_wrap(width = 40) %>% 
  writeLines()
```

* Use geom_hline() and geom_vline() to add reference lines. I often make them thick (size = 2) and white (colour = white), and draw them underneath the primary data layer. That makes them easy to see, without drawing attention away from the data.

* Use geom_rect() to draw a rectangle around points of interest. The boundaries of the rectangle are defined by aesthetics xmin, xmax, ymin, ymax.

* Use geom_segment() with the arrow argument to draw attention to a point with an arrow. Use aesthetics x and y to define the starting location, and xend and yend to define the end location.

# Scales

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete()
```

* the naming scheme for scales
    * scale
    * _
    * the name of the aesthetic
    * _
    * the name of the scale

* the default scale
    * continuous
    * discrete
    * datetime
    * date

## Axis ticks and legend keys

the ticks on the axes.

the keys on the legend.

Breaks controls the position of the ticks, or the values associated with the keys. 

Labels controls the text label associated with each tick/key.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```

Collectively axes and legends are called guides.

*  Axes are used for x and y aesthetics
*  Legends are used for everything else.

```{r}
presidential %>% 
  mutate(id = 33 + row_number()) %>% 
  ggplot(aes(start, id)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y'")
```

* date_labels takes a format specification, in the same form as parse_datetime().

* date_breaks (not shown here), takes a string like “2 days” or “1 month”.

## Legend layout

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  #theme(legend.position = "left")
  #theme(legend.position = "top")
  #theme(legend.position = "bottom")
  theme(legend.position = "right")
  #theme(legend.position = "none")
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))
```

## Replacing a scale

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_color_brewer(palette = "Set1")
  
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_color_brewer(palette = "Set1")
```

```{r}
presidential %>% 
  mutate(id = 33 + row_number()) %>% 
  ggplot(aes(start, id, colour = party)) +
  geom_point() +
  geom_segment(aes(xend = end, yend = id)) +
  scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)

ggplot(df, aes(x, y)) +
  geom_hex() +
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```

Note that all colour scales come in two variety: scale_colour_x( ) and scale_fill_x( ) for the colour and fill aesthetics respectively (the colour scales are available in both UK and US spellings).

# Zooming

There are three ways to control the plot limits: 

1. Adjusting what data are aplotted
2. Setting the limits in each scale
3. Setting xlim and ylim in coord_cartesian( )

```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))
```

```{r}
mpg %>% 
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```

```{r}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point()

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point()
```

```{r}
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_color_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
```

# Themes



```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_bw()
  
```

# Saving your plots

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()
```

```{r}
ggsave("../fig_output/my_plot.pdf")
```

