---
title: "Tidy Data"
author: "Xiaochi"
date: "11/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)

library(tidyverse)
library(nycflights13)
```

# Tidy data

You can represent the same underlying data in multiple ways. The example below shows the same data organised in four different ways. Each dataset shows the same values of four variables country, year, population, and cases, but each dataset organises the values in a different way.

```{r}
table1

table2

table3

table4a

table4b
```

These are all representations of the same underlying data, but they are not equally easy to use. One dataset, the tidy dataset, will be much easier to work with inside the tidyverse.

There are three interrelated rules which make a dataset tidy:

1. Each variable must have its own column.

2. Each observation must have its own row.

3. Each value must have its own cell.

These three rules are interrelated because it’s impossible to only satisfy two of the three. That interrelationship leads to an even simpler set of practical instructions:

1. Put each dataset in a tibble.

2. Put each variable in a column.

In this example, only table1 is tidy. It’s the only representation where each column is a variable.

Why ensure that your data is tidy? There are two main advantages:

1. There’s a general advantage to picking one consistent way of storing data. If you have a consistent data structure, it’s easier to learn the tools that work with it because they have an underlying uniformity.

2. There’s a specific advantage to placing variables in columns because it allows R’s vectorised nature to shine. As you learned in mutate and summary functions, most built-in R functions work with vectors of values. That makes transforming tidy data feel particularly natural.

dplyr, ggplot2, and all the other packages in the tidyverse are designed to work with tidy data. Here are a couple of small examples showing how you might work with table1.

```{r}
# Compute rate per 10,000
table1 %>% 
  mutate(rate=cases/population*10000)

# Compute cases per year
table1 %>% 
  count(year,wt=cases)

# Visualise changes over time
ggplot(table1,aes(year,cases))+
  geom_line(aes(group=country),colour="grey50")+
  geom_point(aes(colour=country))
```

# Spreading and gathering

The principles of tidy data seem so obvious that you might wonder if you’ll ever encounter a dataset that isn’t tidy. Unfortunately, however, most data that you will encounter will be untidy. There are two main reasons:

1. Most people aren’t familiar with the principles of tidy data, and it’s hard to derive them yourself unless you spend a lot of time working with data.

2. Data is often organised to facilitate some use other than analysis. For example, data is often organised to make entry as easy as possible.

This means for most real analyses, you’ll need to do some tidying. 

The first step is always to figure out what the variables and observations are. Sometimes this is easy; other times you’ll need to consult with the people who originally generated the data. 

The second step is to resolve one of two common problems:

1. One variable might be spread across multiple columns.

2. One observation might be scattered across multiple rows.

Typically a dataset will only suffer from one of these problems; it’ll only suffer from both if you’re really unlucky! To fix these problems, you’ll need the two most important functions in tidyr: gather() and spread().

## Gathering

A common problem is a dataset where some of the column names are not names of variables, but values of a variable. Take table4a: the column names 1999 and 2000 represent values of the year variable, and each row represents two observations, not one.

```{r}
table4a
```

To tidy a dataset like this, we need to gather those columns into a new pair of variables. To describe that operation we need three parameters:

* The set of columns that represent values, not variables. In this example, those are the columns 1999 and 2000.

* The name of the variable whose values form the column names. I call that the key, and here it is year.

* The name of the variable whose values are spread over the cells. I call that value, and here it’s the number of cases.

Together those parameters generate the call to gather():

```{r}
table4a %>% 
  gather("1999","2000",key="year",value = "case")
```

The columns to gather are specified with dplyr::select() style notation. Here there are only two columns, so we list them individually. Note that “1999” and “2000” are non-syntactic names (because they don’t start with a letter) so we have to surround them in backticks. To refresh your memory of the other ways to select columns, see select.

n the final result, the gathered columns are dropped, and we get new key and value columns. Otherwise, the relationships between the original variables are preserved. We can use gather() to tidy table4b in a similar fashion. The only difference is the variable stored in the cell values:

```{r}
table4b
```

```{r}
table4b %>% 
  gather("1999","2000",key="year",value="population")
```

To combine the tidied versions of table4a and table4b into a single tibble, we need to use dplyr::left_join(), which you’ll learn about in relational data.

```{r}
tidy4a <- table4a %>% 
  gather("1999","2000",key="year",value="cases")

tidy4b <- table4b %>% 
  gather("1999","2000",key="year",value="population")

left_join(tidy4a,tidy4b)
```


### Select

It’s not uncommon to get datasets with hundreds or even thousands of variables. In this case, the first challenge is often narrowing in on the variables you’re actually interested in. select() allows you to rapidly zoom in on a useful subset using operations based on the names of the variables.

select() is not terribly useful with the flights data because we only have 19 variables, but you can still get the general idea:

```{r}
flights
```

```{r}
# Select columns by name
select(flights,year,month,day)

# Select all columns between year and day (inclusive)
select(flights,year:day)

# Select all columns except those from year to day (inclusive)
select(flights,-(year:day))

# 
```

There are a number of helper functions you can use within select():

* starts_with("abc"): matches names that begin with “abc”.

* ends_with("xyz"): matches names that end with “xyz”.

* contains("ijk"): matches names that contain “ijk”.

* matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.

* num_range("x", 1:3): matches x1, x2 and x3.

select() can be used to rename variables, but it’s rarely useful because it drops all of the variables not explicitly mentioned. Instead, use rename(), which is a variant of select() that keeps all the variables that aren’t explicitly mentioned:

```{r}
rename(flights,tail_num=tailnum)
```

Another option is to use select() in conjunction with the everything() helper. This is useful if you have a handful of variables you’d like to move to the start of the data frame.

```{r}
select(flights,time_hour,air_time,everything())
```

# Spreading

Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. For example, take table2: an observation is a country in a year, but each observation is spread across two rows.

```{r}
table2
```

To tidy this up, we first analyse the representation in similar way to gather(). This time, however, we only need two parameters:

* The column that contains variable names, the key column. Here, it’s type.

* The column that contains values from multiple variables, the value column. Here it’s count.

Once we’ve figured that out, we can use spread(), as shown programmatically below, and visually in Figure 12.3.

```{r}
table2 %>% 
  spread(key=type,value = count)
```

As you might have guessed from the common key and value arguments, spread() and gather() are complements. gather() makes wide tables narrower and longer; spread() makes long tables shorter and wider.

# Separating and uniting

```{r}
table3
```

So far you’ve learned how to tidy table2 and table4, but not table3. table3 has a different problem: we have one column (rate) that contains two variables (cases and population). To fix this problem, we’ll need the separate() function. You’ll also learn about the complement of separate(): unite(), which you use if a single variable is spread across multiple columns.

## Separate

separate() pulls apart one column into multiple columns, by splitting wherever a separator character appears. Take table3:

The rate column contains both cases and population variables, and we need to split it into two variables. separate() takes the name of the column to separate, and the names of the columns to separate into, as shown in Figure 12.4 and the code below.

```{r}
table3 %>% 
  separate(rate,into=c("cases","population"))
```

By default, separate() will split values wherever it sees a non-alphanumeric character (i.e. a character that isn’t a number or letter). For example, in the code above, separate() split the values of rate at the forward slash characters. If you wish to use a specific character to separate a column, you can pass the character to the sep argument of separate(). For example, we could rewrite the code above as:

```{r}
table3 %>% 
  separate(rate,into = c("cases","population"),sep="/")
```

Look carefully at the column types: you’ll notice that cases and population are character columns. This is the default behaviour in separate(): it leaves the type of the column as is. Here, however, it’s not very useful as those really are numbers. We can ask separate() to try and convert to better types using convert = TRUE:

```{r}
table3 %>% 
  separate(rate,into=c("case","population"),convert=TRUE)
```

You can also pass a vector of integers to sep. separate() will interpret the integers as positions to split at. Positive values start at 1 on the far-left of the strings; negative value start at -1 on the far-right of the strings. When using integers to separate strings, the length of sep should be one less than the number of names in into.

You can use this arrangement to separate the last two digits of each year. This make this data less tidy, but is useful in other cases, as you’ll see in a little bit.

```{r}
table3 %>% 
  separate(year,into=c("century","year"),sep=2)
```

## Unite

unite() is the inverse of separate(): it combines multiple columns into a single column. You’ll need it much less frequently than separate(), but it’s still a useful tool to have in your back pocket.

```{r}
table5
```

We can use unite() to rejoin the century and year columns that we created in the last example. That data is saved as tidyr::table5. unite() takes a data frame, the name of the new variable to create, and a set of columns to combine, again specified in dplyr::select() style:

```{r}
table5 %>% 
  unite(new,century,year)
```

In this case we also need to use the sep argument. The default will place an underscore (_) between the values from different columns. Here we don’t want any separator so we use "":

```{r}
table5 %>% 
  unite(new,century,year,sep="")
```

# Missing values

Changing the representation of a dataset brings up an important subtlety of missing values. Surprisingly, a value can be missing in one of two possible ways:

* Explicitly, i.e. flagged with NA.

* Implicitly, i.e. simply not present in the data.

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
```

There are two missing values in this dataset:

* The return for the fourth quarter of 2015 is explicitly missing, because the cell where its value should be instead contains NA.

* The return for the first quarter of 2016 is implicitly missing, because it simply does not appear in the dataset.

One way to think about the difference is with this Zen-like koan: An explicit missing value is the presence of an absence; an implicit missing value is the absence of a presence.

The way that a dataset is represented can make implicit values explicit. For example, we can make the implicit missing value explicit by putting years in the columns:

```{r}
stocks %>% 
  spread(year,return)
```

Because these explicit missing values may not be important in other representations of the data, you can set na.rm = TRUE in gather() to turn explicit missing values implicit:

```{r}
stocks %>% 
  spread(year,return) %>% 
  gather(year,return,"2015":"2016",na.rm = TRUE)
```

Another important tool for making missing values explicit in tidy data is complete():

```{r}
stocks %>% 
  complete(year,qtr)
```

complete() takes a set of columns, and finds all unique combinations. It then ensures the original dataset contains all those values, filling in explicit NAs where necessary.

There’s one other important tool that you should know for working with missing values. Sometimes when a data source has primarily been used for data entry, missing values indicate that the previous value should be carried forward:

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
```

You can fill in these missing values with fill(). It takes a set of columns where you want missing values to be replaced by the most recent non-missing value (sometimes called last observation carried forward).

```{r}
treatment %>% 
  fill(person)
```

## Case Study

To finish off the chapter, let’s pull together everything you’ve learned to tackle a realistic data tidying problem. The tidyr::who dataset contains tuberculosis (TB) cases broken down by year, country, age, gender, and diagnosis method. The data comes from the 2014 World Health Organization Global Tuberculosis Report, available at http://www.who.int/tb/country/data/download/en/.

There’s a wealth of epidemiological information in this dataset, but it’s challenging to work with the data in the form that it’s provided:

```{r}
who
```

This is a very typical real-life example dataset. It contains redundant columns, odd variable codes, and many missing values. In short, who is messy, and we’ll need multiple steps to tidy it. Like dplyr, tidyr is designed so that each function does one thing well. That means in real-life situations you’ll usually need to string together multiple verbs into a pipeline.

The best place to start is almost always to gather together the columns that are not variables. Let’s have a look at what we’ve got:

* It looks like country, iso2, and iso3 are three variables that redundantly specify the country.

* year is clearly also a variable.

* We don’t know what all the other columns are yet, but given the structure in the variable names (e.g. new_sp_m014, new_ep_m014, new_ep_f014) these are likely to be values, not variables.

So we need to gather together all the columns from new_sp_m014 to newrel_f65. We don’t know what those values represent yet, so we’ll give them the generic name "key". We know the cells represent the count of cases, so we’ll use the variable cases. There are a lot of missing values in the current representation, so for now we’ll use na.rm just so we can focus on the values that are present.

```{r}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65,key = "key",value="cases",na.rm=TRUE)
```

We can get some hint of the structure of the values in the new key column by counting them:

```{r}
who1 %>% 
  count(key)
```

You might be able to parse this out by yourself with a little thought and some experimentation, but luckily we have the data dictionary handy. It tells us:

1. The first three letters of each column denote whether the column contains new or old cases of TB. In this dataset, each column contains new cases

2. The next two letters describe the type of TB:
  + rel stands for cases of relapse
  + stands for cases of extrapulmonary TB
  + sn stands for cases of pulmonary TB that could not be diagnosed by a pulmonary smear (smear negative)
  + sp stands for cases of pulmonary TB that could be diagnosed be a pulmonary smear (smear positive)

3. The sixth letter gives the sex of TB patients. The dataset groups cases by males (m) and females (f).

4. The remaining numbers gives the age group. The dataset groups cases into seven age groups:
  + 014 = 0 – 14 years old
  + 1524 = 15 – 24 years old
  + 2534 = 25 – 34 years old
  + 3544 = 35 – 44 years old
  + 4554 = 45 – 54 years old
  + 5564 = 55 – 64 years old
  + 65 = 65 or older

We need to make a minor fix to the format of the column names: unfortunately the names are slightly inconsistent because instead of new_rel we have newrel (it’s hard to spot this here but if you don’t fix it we’ll get errors in subsequent steps). You’ll learn about str_replace() in strings, but the basic idea is pretty simple: replace the characters “newrel” with “new_rel”. This makes all variable names consistent.

```{r}
who2 <- who1 %>% 
  mutate(key=stringr::str_replace(key,"newrel","new_rel"))
```

We can separate the values in each code with two passes of separate(). The first pass will split the codes at each underscore.

```{r}
who3 <- who2 %>% 
  separate(key,c("new","type","sexage"),sep="_")
```

Then we might as well drop the new column because it’s constant in this dataset. While we’re dropping columns, let’s also drop iso2 and iso3 since they’re redundant.

```{r}
who3 %>% 
  count(new)
```

```{r}
who4 <- who3 %>% 
  select(-new,-iso2,-iso3)
```

Next we’ll separate sexage into sex and age by splitting after the first character:

```{r}
who5 <- who4 %>% 
  separate(sexage,c("sex","age"),sep = 1)
```

The who dataset is now tidy!

I’ve shown you the code a piece at a time, assigning each interim result to a new variable. This typically isn’t how you’d work interactively. Instead, you’d gradually build up a complex pipe:

```{r}
who %>% 
  gather(key,value,new_sp_m014:newrel_f65,na.rm=TRUE) %>% 
  mutate(key=stringr::str_replace(key,"newrel","new_rel")) %>% 
  separate(key,c("new","var","sexage")) %>% 
  select(-new,-iso2,-iso3) %>% 
  separate(sexage,c("sex","age"),sep=1)
```
