---
title: "Workflow"
author: "Xiaochi"
date: "05/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

object name gets value

I recommend snake_case where you separate lowercase words with _.

```{r}
this_is_a_really_long_name <- 2.5

this_is_a_really_long_name

this

this
```

You will make lots of assignments and <- is a pain to type. Don’t be lazy and use =: it will work, but it will cause confusion later. Instead, use RStudio’s keyboard shortcut: Alt + - (the minus sign). 

Ooops, you made a mistake! this_is_a_really_long_name should have value 3.5 not 2.5. Use another keyboard shortcut to help you fix it. Type “this” then press Cmd/Ctrl + ↑. That will list all the commands you’ve typed that start those letters. Use the arrow keys to navigate, then press enter to retype the command. Change 2.5 to 3.5 and rerun.(only in console)

If you make an assignment, you don’t get to see the value. You’re then tempted to immediately double-check the result:

This common action can be shortened by surrounding the assignment with parentheses, which causes assignment and “print to screen” to happen.

Press Alt + Shift + K