---
title: "ResBaz"
author: "Xiaochi"
date: "10/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

# Day 1

## Python

Python gets its name from the "Monty Pthon's Flying Circus" comedy series.

```{python}
import numpy as np
import matplotlib as plt
```

## R

R. Ihaka and R. Gentleman

Designed for statistical computing and graphics

* Tidyverse
    * readr (data import)
    * dplyr, tidyr (data manipulation)
    * ggplot2 (visualisation)
* zoo (time series)
* RTextTools (machine learning for text classification)
* caret (machine learning)

## MATLAB

technical computing platform designed for Engineers and Scientists

## Julia

A Fast Dynamic Language for Technical Computing

2012, MIT

juliacomputing.com

Open scource

Efficient for program that use loops

Enables domain experts to produce code that has performance suitable for production use - no rework or code porting to a compiled language is needed to get this.

Ability to call C, Python functions

## Workflow

## Publishing

Targeting your audience

Bibliometrics
Three levels: author, article and journal
Concerns: how research has influenced another reseracher or research organisation
Measures include: citation counts, citation patterns

* find journal
    * JANE
    * Joun





# Day 2

## Raster

pixel depth
resolution

LiDAR
Landsat

## Vector

## CRS

Datum

Projection
Universal transverse Mercator (UTM)

tmap, leaflet

## Live Code

```{r}
library(sp)
library(raster)
library(rgdal)
library(tidyverse)
```

```{r}
GDALinfo("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
```

```{r}
HARV_dsmCrop_info <- capture.output(
    GDALinfo("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
    )
```

```{r}
DSM_HARV <- raster("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
```

```{r}
summary(DSM_HARV)
```

```{r}
summary(DSM_HARV, maxsamp = ncell(DSM_HARV))
```

```{r}
DSM_HARV_df <- as.data.frame(DSM_HARV, xy = TRUE)
str(DSM_HARV_df)
```

```{r}
ggplot() +
    geom_raster(data = DSM_HARV_df, 
                aes(x = x, y = y, fill = HARV_dsmCrop)) +
    scale_fill_viridis_c() +
    coord_quickmap()
```

```{r}
crs(DSM_HARV)
minValue(DSM_HARV)
maxValue(DSM_HARV)
```

```{r}
DSM_HARV <- setMinMax(DSM_HARV)
nlayers(DSM_HARV)
```

What units are our data in?
```{r}
crs(DSM_HARV)
```

```{r}
DSM_HARV
```


Use the output from the GDALinfo() function to find out what NoDataValue is used for our DSM_HARV dataset.
```{r}
GDALinfo("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
```

```{r}
ggplot() +
    geom_histogram(data = DSM_HARV_df, aes(HARV_dsmCrop))
```

```{r}
DSM_HARV_df <- DSM_HARV_df %>% 
    mutate(fct_elevation = cut(HARV_dsmCrop, breaks = 3))
str(DSM_HARV_df)
```

```{r}
ggplot() +
    geom_bar(data = DSM_HARV_df, aes(fct_elevation))
```

```{r}
unique(DSM_HARV_df$fct_elevation)
```

```{r}
DSM_HARV_df %>% 
    group_by(fct_elevation) %>% 
    count()
```

```{r}
custom_bins <- c(300, 350, 400, 450)
DSM_HARV_df <- DSM_HARV_df %>% 
    mutate(fct_elevation_2 = cut(HARV_dsmCrop, breaks = custom_bins))
str(DSM_HARV_df)
unique(DSM_HARV_df$fct_elevation_2)
```

```{r}
ggplot() +
    geom_bar(data = DSM_HARV_df, aes(fct_elevation_2))
```

```{r}
ggplot() +
    geom_raster(data = DSM_HARV_df, 
                aes(x = x, y = y, fill = fct_elevation_2)) +
    scale_fill_manual(values = terrain.colors(3)) +
    coord_quickmap()
```

Challenge 4
Create a plot of the Harvard Forest Digital Surface Model (DSM) that has:
1. Six classified ranges of values (break points) that are evenly divided among the range of pixel values.
2. Axis labels.
3. A plot title.

```{r}
DSM_HARV_df <- DSM_HARV_df %>% 
    mutate(fct_elevation_3 = cut(HARV_dsmCrop, breaks = 6))
unique(DSM_HARV_df$fct_elevation_3)
```


```{r}
ggplot() +
    geom_raster(data = DSM_HARV_df, 
                aes(x = x, y = y, fill = fct_elevation_3)) +
    scale_fill_manual(values = terrain.colors(6)) +
    labs(title = "Harvard forest elevation(m)",
         x = "projection coordinate x",
         y = "projection coordinate y") +
    coord_quickmap()
```

```{r}
GDALinfo("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_DSMhill.tif")

```

```{r}
crs(DSM_HARV)
```


## Skills for 21st century researchers

## Intersect

Accelerate your research

Productivity
Diversity
Longevity

dashboard quick visualisation

## German
huong-ly.tong@students.mq.edu.au
Huong Ly Tong

tiny.cc/ds3goe

# Day 3

## Raster

DTM: elevation

```{r}
DTM_HARV <- raster("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DTM/HARV_dtmCrop.tif")

DTM_hill_HARV <- raster("../data/NEON-DS-Airborne-Remote-Sensing/HARV/DTM/HARV_DTMhill_WGS84.tif")

DTM_HARV_df <- as.data.frame(DTM_HARV, xy = TRUE)
DTM_hill_HARV_df <- as.data.frame(DTM_hill_HARV, xy = TRUE)
```


```{r}
DTM_HARV_df %>% 
    ggplot(aes(x = x, y = y, fill = HARV_dtmCrop)) +
    geom_raster() +
    scale_fill_gradientn(name = "Elevation", 
                        colours = terrain.colors(10)) +
    coord_quickmap()
```

```{r}
DTM_hill_HARV_df %>% 
    ggplot(aes(x = x, y = y, alpha = HARV_DTMhill_WGS84)) +
    geom_raster() +
    coord_quickmap()
```

change porjection

```{r}
DTM_hill_UTMZ18_HARV <- projectRaster(DTM_hill_HARV,
                                      crs = crs(DTM_HARV))
crs(DTM_hill_UTMZ18_HARV)
crs(DTM_HARV)
extent(DTM_hill_UTMZ18_HARV)
extent(DTM_HARV)
res(DTM_hill_UTMZ18_HARV)

# force resolution
DTM_hill_UTMZ18_HARV <- projectRaster(DTM_hill_HARV,
                                      crs = crs(DTM_HARV),
                                      res = 1)
res(DTM_hill_UTMZ18_HARV)

DTM_hill_UTMZ18_HARV_df <- as.data.frame((DTM_hill_UTMZ18_HARV), xy = TRUE)
```


```{r}
ggplot() +
    geom_raster(data = DTM_hill_UTMZ18_HARV_df,
                aes(x = x, y = y, alpha = HARV_DTMhill_WGS84)) +
    geom_raster(data = DTM_HARV_df,
                aes(x = x, y = y, fill = HARV_dtmCrop),
                alpha = 0.3) +
    scale_fill_gradientn(colours = terrain.colors(10)) +
    coord_quickmap()
```


```{r}
DSM_HARV
DTM_HARV

CHM_HARV <- DSM_HARV - DTM_HARV
```

quick map

```{r}
plot(CHM_HARV)
```

```{r}
CHM_HARV_df <- as.data.frame(CHM_HARV, xy = TRUE)
ggplot() +
    geom_histogram(data = CHM_HARV_df,
                   aes(layer))
```

```{r}
CHM_ov_HARV <- overlay(DSM_HARV, 
                       DTM_HARV, 
                       fun = function(r1, r2) {
                           return(r1 - r2)
                       })
plot(CHM_ov_HARV)

CHM_ov_HARV_df <- as.data.frame(CHM_ov_HARV, xy = TRUE)

ggplot() +
    geom_histogram(data = CHM_ov_HARV_df,
                   aes(layer))

```

write out

```{r}
writeRaster(CHM_ov_HARV,
            "../data/NEON-DS-Airborne-Remote-Sensing/HARV/CHM/CHM_HARV.tif",
            format = "GTiff",
            overwrite = TRUE,
            NAflag = -9999)
```

## Vector

```{r}
library(sf)
```

```{r}
aoi_boundary_HARV <- st_read("../data/NEON-DS-Site-Layout-Files/HARV/HarClip_UTMZ18.shp")
```

```{r}
st_geometry_type(aoi_boundary_HARV)
levels(st_geometry_type(aoi_boundary_HARV))

st_crs(aoi_boundary_HARV)
# extent
st_bbox(aoi_boundary_HARV)

theme_set(theme_minimal())

aoi_boundary_HARV %>% 
    ggplot() +
    geom_sf(size = 3, color = "black", fill = "cyan1")
```

```{r}
lines_HARV <- st_read("../data/NEON-DS-Site-Layout-Files/HARV/HARV_roads.shp")
point_HARV <- st_read("../data/NEON-DS-Site-Layout-Files/HARV/HARVtower_UTM18N.shp")

st_crs(lines_HARV)
st_crs(point_HARV)

st_bbox(lines_HARV)
st_bbox(point_HARV)

lines_HARV %>% head()
```

Explore the attributes associated with the point_HARV and aoi_boundary_HARV spatial objects.
1. How many attributes does each have?
```{r}
ncol(point_HARV)
ncol(aoi_boundary_HARV)
```

2. Who owns the site in the point_HARV data object?
```{r}
point_HARV$Ownership
```

```{r}
point_HARV["Ownership"]
```

3. Which of the following is NOT an attribute of the point_HARV data object?
C) Country

```{r}
lines_HARV %>% 
    ggplot() +
    geom_sf(aes(col = TYPE)) +
    scale_color_manual(values = c("red", "blue", "green", "magenta")) +
    ggtitle("HARV Forest", subtitle = "Roads by taype") +
    coord_sf()
```

Challenge 14
1. Create a map of the state boundaries in the United States using the data located in your downloaded data folder: 
NEON-DS-Site-Layout-Files/US-Boundary-Layers\US-State-Boundaries-Census-2014.

```{r}
state_boundary_us <- st_read("../data/NEON-DS-Site-Layout-Files/US-Boundary-Layers/US-State-Boundaries-Census-2014.shp")
```

```{r}
ncol(state_boundary_us)
```


```{r}
state_boundary_us %>% 
    ggplot() +
    geom_sf() +
    coord_sf()
```

2. Apply a fill color to each state using its region value. 
```{r}
state_boundary %>% 
    ggplot() +
    geom_sf(aes(fill = region)) +
    coord_sf()
```

## Why CRS

Challenge 15
Create a map of the North Eastern United States as follows:
1. Import and plot Boundary-US-State-NEast.shp. Adjust line width as necessary.
```{r}
boundary_us <- st_read("../data/NEON-DS-Site-Layout-Files/US-Boundary-Layers/Boundary-US-State-NEast.shp")

point_HARV_4326 <- st_transform(point_HARV, st_crs(boundary_us))

ggplot() +
    geom_sf(data = boundary_us,
            aes(fill = STUSPS)) +
    scale_fill_brewer(palette = "Set3") +
    geom_sf(data = point_HARV_4326,
            aes(shape = "shape")) +
    labs(title = "Havard tower") +
    coord_sf()
```

2. Layer the Fisher Tower (in the NEON Harvard Forest site) point location point_HARV onto the plot.
3. Add a title.
4. Add a legend that shows both the state boundary (as a line) and the Tower location point.



find the story
find multiple stories

## communication

start with a personal anecdote...

Not too much text
Don't read the slides
Put the presentation in fullscreen mode
Don't point at your laptop
Practice

# Compute Day

Access and Explore
Preprocess Data

Data Analysis/Analytics
Share/Deploy


Machine Learning
Deep Learning
Reinforcement Learning

which is the best network

preallocation


# Research Computing

Macquarie Subscription
my.nci.org.au

www.infrastructure.aws

rstudio.science.mq.edu.au

hive.science.mq.edu.au/hpc

NCI

mq-data-advice.talkyard.net

